### 2022-03-11 Report/NO2.Rmd
Upload New File

---
title: "NO2  case study (5 stations)"
author: "HW"
date: "March 12, 2019"
output: html_document
---

Data set from the paper  by Cooley,  Davis, Naveau (The Annals of Applied Statistics 2012, Vol. 6, No. 4, 1406–1429). 
 Data available  at  http://www.stat.colostate.edu/~cooleyd/DataAndCode/PredExtremes/.

## Description

- Variable:  daily maximum hourly NO2 measurement.

- Five stations located in Washington DC and nearby Virginia: Alexandria (alx), McMillan (mc), River Terrace (rt), Takoma School (ts) and Arlington (arl).

- Period: 5000 daily NO2 measurements recorded between 1995 and 2010, keeping only those records for which all five stations have measurements, resulting in 4497 daily measurements.

- Rounding: because the data were truncated to the nearest ppb, a uniform random variable on the interval [−0.5, 0.5] was added to the data so that they behave more like the underlying continuous variable.

```{r}

## Load data and display first and last 5 lines

load("nitrogenDioxideMsmts.RData")
head(no2Msmts); tail(no2Msmts)
```

```{r}

## Load graphical package lattice 
library(lattice)

## Times series at each of the 5 stations
dates=as.Date(no2Msmts[,1], origin = "1970-01-01")
xyplot(no2Msmts[,2]~ dates,pch=19,ylab="NO2 at Alexandria")
xyplot(no2Msmts[,3]~ dates,pch=19,ylab="NO2 at Arlington")
xyplot(no2Msmts[,4]~ dates,pch=19,ylab="NO2 at McMillan ")
xyplot(no2Msmts[,5]~ dates,pch=19,ylab="NO2 at River Terrace")
xyplot(no2Msmts[,6]~ dates,pch=19,ylab="NO2 at Takoma School")
```

## Examine dependence on weekdays

On Saturday and Sunday there is less NO2 for all series.

```{r}
## data object is a matrix, create a data frame
class(no2Msmts)
NO2=as.data.frame(no2Msmts)
names(NO2)
```

```{r}

## in case the locale is not in english
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

```{r}

## dependence on weekdays: on Sat and Sun there is decrease of NO2
day= factor( weekdays(dates,abbreviate=TRUE), levels=c("Mon","Tue","Wed","Thu", "Fri", "Sat", "Sun"),ordered=TRUE)

bwplot(day~NO2$ts, main="Tacoma school station",xlab="NO2")

```

## Examine dependence on month

Strangely the median is highest for Feb, Mar, Apr for all series

```{r}
month =factor( months(dates,abbreviate=TRUE),levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE)

bwplot(month~NO2$alx);bwplot(month~NO2$arl);bwplot(month~NO2$mc);bwplot(month~NO2$rt);bwplot(month~NO2$ts)

```

## QQ plots

Empirical quantiles vs normal quantiles

```{r}
qqmath(NO2$alx[month=="Jul"])
```

Empirical quantiles vs exponential quantiles

```{r}
qqmath(NO2$alx[month=="Jul"],distri=qexp)
```

## Extreme value distribution fit

We use the library ismev that goes with the book by Stuart Coles

```{r}
# the library ismev needs to be downloaded from CRAN to your machine
# install.packages("ismev")

## Load library ismev
library(ismev)
```

### GEV modelling for each station

Modelling all data without distinguishing days and months, the tail is not fitted well at all for each of the series.

```{r]
## all days and months together
gev.diag(gev.fit(NO2$alx));gev.diag(gev.fit(NO2$arl));gev.diag(gev.fit(NO2$mc));gev.diag(gev.fit(NO2$rt));gev.diag(gev.fit(NO2$ts))
```

### GEV modelling performed for each month

Better fits are obtained for individual months

```{r}
## for a specific month
gev.diag(gev.fit(NO2$alx[month=="Jul"]));gev.diag(gev.fit(NO2$arl[month=="Jul"]));gev.diag(gev.fit(NO2$mc[month=="Jul"]));gev.diag(gev.fit(NO2$rt[month=="Jul"]));gev.diag(gev.fit(NO2$ts[month=="Jul"]))

```

## Mean excess plot

```{r}
mrl.plot(NO2$alx[month=="Jul"])
```

## Scatter plot matrix for the 5 stations

No distinction of months.

Observe the extremes: are they common for a pair of stations ?

```{r}
splom(NO2[,2:6],pch=19)

```

## Scatter plot matrix for the 5 stations

Using a different colour for each month.

```{r}
splom(NO2[,2:6],group=as.factor(month),auto.key=TRUE)
```

### 2022-03-11 install.R
Update install.R

@@ -2,3 +2,30 @@ install.packages("evd")
 install.packages("evir")
 install.packages("ismev")
 install.packages("SpatialExtremes")
+
+# Update markdown pacakge
+install.packages("markdown")
+
+# Packages for plotting spatial objects
+install.packages("RgoogleMaps")
+install.packages("png")
+install.packages("rjson")
+install.packages("jpeg")
+install.packages("bitops")
+install.packages("plogr")
+
+install.packages("sp")
+install.packages("rgdal")
+install.packages("ggplot2")
+install.packages("evd")
+install.packages("gmp")
+install.packages("mvPot")
+
+install.packages("git2r")
+
+install.packages("usethis")
+install.packages("devtools")
+#devtools::install_github("hadley/devtools")
+devtools::install_github("dkahle/ggmap")
+
+

### 2022-03-11 Practicals/Day1_DJIA.R
Upload New File

@@ -1,0 +1,700 @@
+DJIA <-
+list(structure(c(15043, 15044, 15047, 15048, 15049, 15050, 15051, 
+15054, 15055, 15056, 15057, 15058, 15061, 15062, 15063, 15064, 
+15065, 15068, 15069, 15070, 15071, 15072, 15075, 15076, 15077, 
+15078, 15079, 15082, 15083, 15084, 15085, 15089, 15090, 15091, 
+15092, 15093, 15096, 15097, 15098, 15099, 15100, 15103, 15104, 
+15105, 15106, 15107, 15110, 15111, 15112, 15113, 15114, 15117, 
+15118, 15119, 15120, 15121, 15125, 15126, 15127, 15128, 15131, 
+15132, 15133, 15134, 15135, 15138, 15139, 15140, 15141, 15142, 
+15145, 15146, 15147, 15148, 15149, 15152, 15153, 15154, 15155, 
+15156, 15160, 15161, 15162, 15163, 15166, 15167, 15168, 15169, 
+15170, 15173, 15174, 15175, 15176, 15177, 15180, 15181, 15182, 
+15183, 15184, 15187, 15188, 15189, 15190, 15191, 15194, 15195, 
+15196, 15197, 15198, 15201, 15202, 15203, 15204, 15205, 15208, 
+15209, 15210, 15211, 15212, 15215, 15216, 15217, 15218, 15219, 
+15223, 15224, 15225, 15226, 15229, 15230, 15231, 15232, 15233, 
+15236, 15237, 15238, 15239, 15240, 15243, 15244, 15245, 15246, 
+15247, 15250, 15251, 15252, 15253, 15254, 15257, 15258, 15259, 
+15260, 15261, 15264, 15265, 15266, 15267, 15268, 15271, 15272, 
+15273, 15274, 15275, 15278, 15279, 15280, 15281, 15282, 15285, 
+15286, 15287, 15288, 15289, 15292, 15293, 15294, 15295, 15296, 
+15299, 15300, 15301, 15303, 15306, 15307, 15308, 15309, 15310, 
+15313, 15314, 15315, 15316, 15317, 15320, 15321, 15322, 15323, 
+15324, 15327, 15328, 15329, 15330, 15331, 15335, 15336, 15337, 
+15338, 15342, 15343, 15344, 15345, 15348, 15349, 15350, 15351, 
+15352, 15356, 15357, 15358, 15359, 15362, 15363, 15364, 15365, 
[…]

### 2022-03-11 Practicals/Day1_MEUSE.Rmd
Upload New File

---
title: "Meuse river annual maximal discharge"
author: "HW"
date: "May 24, 2018"
output: html_document
---

Meuse annual maximal discharge
==

```{r, echo=FALSE}
# LIBRARY: Intro Stat Modeling Extreme Values
library(ismev)
```

```{r}
# import meuse data
source("meuse.data.R")
attach(meuse)
require(lattice)
names(meuse)
xyplot(Discharge~Year,pch=19)
```

Meuse annual maximal discharge
==

```{r}

dotplot(Discharge)
bwplot(Discharge)
histogram(Discharge,n=20)
```

Normal and exponential QQ plots
==

```{r}
qqmath(Discharge,xlab="Normal quantiles")
# QQplot for Extreme Value Analysis
qqmath(Discharge,distribution = qexp,xlab="Exponential quantiles")

```

Two versions of the excess plot
==

```{r}
# evir library: excess plot
library(evir)
meplot(Discharge)
# ismev library: excess plot
mrl.plot(Discharge)
# notice that meplot has a parameter omit=3: three largest values are omitted
?meplot

```

GEV and Gumbel model fits
=

```{r}
meuse.GEV=gev.fit(Discharge)
meuse.GUM=gum.fit(Discharge)
# plot diagnostics
gev.diag(meuse.GEV)
gum.diag(meuse.GUM)
```
```{r}
gev.profxi(meuse.GEV)   #-.3,.3
gev.prof(meuse.GEV,100) # 2500,4500
```
CONCLUSION
=

Same conclusion as Beirlant et al.:
Gumbel model is better than GEV model.

NEW ANALYSIS OF THAT DATA
==

Discussion: are there outliers?

```{r}
plot(meuse)
lines(lowess(meuse),col=2)

```

Which years?
==

```{r}
# there seems to be an outlier: which year?
plot(meuse)
###  For interactivity execute the commands of chunk in the CONSOLE
# CLICK on point in Plot window [ESC  to end]
identify(meuse)
```

Outliers
==

```{r}
Year[15]; Year[85]

```
Remove outliers
==

```{r}
mm2 = data.frame(cbind(Year[c(-15,-85)],Discharge[c(-15,-85)]))
colnames(mm2) = c("Year","Discharge")
```

Recompute and display the overall trend without outlier
==

```{r}
plot(meuse) # with outlier
# using "loess"" instead of "lowess""
lines(meuse$Year,
predict(loess(Discharge ~ Year,
data=meuse,span=1.2)) ,col=4)

plot(mm2) # without outlier
lines(mm2$Year,
predict(loess(Discharge ~ Year,
data=mm2,span=1.2)) ,col=2)

```

Including the trend into the GEV analysis
==

```{r}
## storing the trend
meuse.loess =
as.data.frame(
predict(loess(Discharge ~ Year,
data=mm2,span=1.2)))
colnames(meuse.loess)="CE"

## storing the analysis

meuse.GEVloess =
gev.fit(mm2$Discharge,ydat=meuse.loess,mul=1)
meuse.GUMloess =
gum.fit(mm2$Discharge,ydat=meuse.loess,mul=1)

# plot diagnostics
gev.diag(meuse.GEVloess)
gum.diag(meuse.GUMloess)

```
```{r}
## compare the negative log-likelihoods (nllh)
meuse.GEVloess$nllh
meuse.GUM$nllh
```

CONCLUSION
==

GEV model fits the residuals better than Gumbel model.

### 2022-03-11 Code_examples/plot_rainfall.R
Add code plots.

@@ -1,0 +1,20 @@
+load("team_number.RData")
+
+library(ggmap)
+bbox <- c(141.37838, -39.21238, 145.35653, -36.10332)
+map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")
+
+dataToPlot = data.frame(X = coordinates@coords[, 1], Y = coordinates@coords[, 2])
+mapPoints <- ggmap(map) +
+  geom_point(
+    data = dataToPlot,
+    aes(x = X, y = Y),
+    color = "red",
+    alpha = 1,
+    size = 3,
+    shape = 16,
+  ) +
+  labs(title = "", fill = "") +
+  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
+  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
+print(mapPoints)

### 2022-03-11 Code_examples/plot_temperature.R

@@ -1,0 +1,19 @@
+load("team_number.RData")
+
+bbox <- c(-8.625, 38.375, 14.625, 49.625)
+map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")
+
+dataToPlot = data.frame(X = coordinates@coords[, 1], Y = coordinates@coords[, 2])
+mapPoints <- ggmap(map) +
+  geom_point(
+    data = dataToPlot,
+    aes(x = X, y = Y),
+    color = "red",
+    alpha = 1,
+    size = 3,
+    shape = 16,
+  ) +
+  labs(title = "", fill = "") +
+  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
+  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
+print(mapPoints)

### 2022-03-11 Code_examples/plot_wind.R

@@ -1,0 +1,19 @@
+load("team_number.RData")
+
+bbox <- c(-21.375,  40.875,  16.875,  61.125)
+map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")
+
+dataToPlot = data.frame(X = coordinates[, 1], Y = coordinates[, 2])
+mapPoints <- ggmap(map) +
+  geom_point(
+    data = dataToPlot,
+    aes(x = X, y = Y),
+    color = "red",
+    alpha = 1,
+    size = 3,
+    shape = 16,
+  ) +
+  labs(title = "", fill = "") +
+  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
+  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
+print(mapPoints)

### 2022-03-14 install.R
Update install.R with package for Thomas.

@@ -2,6 +2,11 @@ install.packages("evd")
 install.packages("evir")
 install.packages("ismev")
 install.packages("SpatialExtremes")
+install.packages("BMAmevt")
+install.packages("extRemes")
+install.packages("fExtremes")
+install.packages("mev")
+install.packages("MCMCpack")
 
 # Update markdown pacakge
 install.packages("markdown")

### 2022-03-15 Practicals/multivariate2022.Rmd
Upload New File

---
title: "ATHENS course on Extreme-Value Theory"
subtitle:  "Multivariate extreme-value analysis"
author: "Thomas Opitz (BioSP, INRAE)"
date: "16 March 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, tidy = FALSE, dev = 'png')
par(mar = c(5, 4, 1, 1) + .1)
WORK = "~/research/enseignement/ATHENS_Cours-Extremes/code/"
setwd(WORK)
options(warn=-1)
```

##Practicals Part 1

### 1.1) Simulation and visualisation of the pairwise beta angular measure

We sample from the spectral measure using the function rpairbeta, and then we provide an interactive 3D plot with plot3d. 
We can also add the boundary segments of the unit simplex, and the origin (0,0,0).

First, we load the required R packages. 
```{r load packages, message=FALSE}
#library containing the pairwise beta angular measure:
library(BMAmevt)
#library for "3D Visualization Using OpenGL":
library(rgl)
```

```{r, echo=TRUE}
n=200
pars=c(0.5,0.2,0.2,6)
w=rpairbeta(n,dimData=3,par=pars)
plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))

pars=c(.1,10,10,10)
w=rpairbeta(n,dimData=3,par=pars)
plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))
segments3d(c(1,0),c(0,1),c(0,0),col="blue")
segments3d(c(1,0),c(0,0),c(0,1),col="blue")
segments3d(c(0,0),c(0,1),c(1,0),col="blue")
points3d(0,0,0,col="blue",size=10)

pars=c(25,10,10,10)
w=rpairbeta(n,dimData=3,par=pars)
plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))
segments3d(c(1,0),c(0,1),c(0,0),col="blue")
segments3d(c(1,0),c(0,0),c(0,1),col="blue")
segments3d(c(0,0),c(0,1),c(1,0),col="blue")
points3d(0,0,0,col="blue",size=10)

pars=c(.1,.1,.1,.1)
w=rpairbeta(n,dimData=3,par=pars)
plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))
segments3d(c(1,0),c(0,1),c(0,0),col="blue")
segments3d(c(1,0),c(0,0),c(0,1),col="blue")
segments3d(c(0,0),c(0,1),c(1,0),col="blue")
points3d(0,0,0,col="blue",size=10)

pars=c(25,.1,.1,.1)
w=rpairbeta(n,dimData=3,par=pars)
plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))
segments3d(c(1,0),c(0,1),c(0,0),col="blue")
segments3d(c(1,0),c(0,0),c(0,1),col="blue")
segments3d(c(0,0),c(0,1),c(1,0),col="blue")
points3d(0,0,0,col="blue",size=10)
```

### 1.2) Empirical estimation of an extreme event probability

We first define the function ell.min that appropriately reweights the angles of w. Then, we can empirically estimate the target probabilities.

```{r, echo=TRUE}
ell.min=function(w,u){min(w/u)}
pars=c(5,1,1,1)
w=rpairbeta(n,dimData=3,par=pars)
3/n*sum(apply(w,1,ell.min,u=10))
pars=c(.1,1,1,1)
w=rpairbeta(n,dimData=3,par=pars)
3/n*sum(apply(w,1,ell.min,u=10))
```
## Practicals Part 2: Dependence summaries of air pollution data

### Description (recall of Day 1 practicals)

- Variable:  daily maximum hourly NO2 measurement.

- Five observation stations located in Washington DC and nearby  Virginia: Alexandria (alx), McMillan (mc), River Terrace (rt), Takoma School (ts) and Arlington (arl).

- Observation period: 5000 daily NO2 measurements recorded between 1995 and 2010, where only those records are kept for which all five stations have measurements, resulting in 4497 daily measurements.

- Rounding: because the data were truncated to the nearest ppb value, a uniform random variable on the interval (-0.5, 0.5) was added to the data so that they behave more like the underlying continuous variable.

We first load the data object and display its first and last lines.

```{r, echo=TRUE}
load("nitrogenDioxideMsmts.RData")
dim(no2Msmts)
head(no2Msmts)
tail(no2Msmts)
```

For simplicity, we will ignore seasonal and long-term trends in the time series and proceed as if data vectors were independent and identically distributed over time. We extract the five data columns into a matrix, and then we show pairwise scatterplots.

```{r}
mydata=no2Msmts[,-1]
pairs(mydata)
```

### 2.1) Marginal standardization based on the rank transform

We now transform the data to standardized margins (standard Pareto here) using the empirical probability distribution, that is, we use the so-called rank transform.

```{r, echo=TRUE}
n=nrow(mydata)
myranks=apply(mydata,2,rank)
data.unif=myranks/(n+1)
data.stand=1/(1-data.unif)
```

### 2.2) Estimation of the extremal coefficient

We estimate the extremal coefficient for a range of values. To obtain symmetric confidence intervals, we use a simple estimator of the standard deviation (that disregards the uncertainty coming from the marginal pretransformation to the standard Pareto scale).

```{r, echo=TRUE}
p=0.8+0:19*0.01
u.star=1/(1-p)
u.star[1]*mean(apply(data.stand>u.star[1],1,any))
u.star[20]*mean(apply(data.stand>u.star[20],1,any))
theta.est=c()
for(i in 1:length(p)){
  theta.est[i]=u.star[i]*mean(apply(data.stand>u.star[i],1,any))
}
plot(p,theta.est,type="l",lwd=3,xlab="p",ylab=expression(theta[5]))
# calculate approximate 95% pointwise confidence intervals
sd_est = sqrt(u.star/n * theta.est *(1-theta.est/u.star))
lines(p,theta.est + qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
lines(p,theta.est - qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
```

Judging from the significantly increasing estimates when p increases, the data do not show threshold-stability. We seem to be in the situation of asymptotic independence.

### 2.3) Estimation of the coefficient of tail dependence

Next, we plot the estimates of the coefficient of tail dependence, estimated using the Hill estimator applied to the sample of the minimum of the 5 components. This can be done conveniently with the hillPlot function in the fExtremes package, which also provides confidence intervals.

```{r load packages 2, message=FALSE}
library(fExtremes)
```

```{r, echo=TRUE}
hillPlot(apply(data.stand,1,min), start=0.8*n, plottype="xi") 
```

The values estimated for eta, and the upper confidence bound far below 1, strongly suggest that we have joint asymptotic independence for these 5 variables.

## Practicals Part 3: Modeling extremal dependence in air pollution data

### 3.1) Empirical estimation of the angular measure

We first construct the pseudo-polar coordinates and then extract the angles w corresponding to extreme events.

```{r, echo=TRUE}
r.obs=apply(data.stand,1,sum)
w.obs=data.stand/r.obs
r_0=quantile(r.obs,0.95)
w.H=w.obs[r.obs>r_0,]
```

### 3.2) Estimation of joint exceedances above the overall 95% quantile

```{r, echo=TRUE}
q95=quantile(mydata,0.95)
q95
prob.component=apply(mydata<q95,2,mean)
u.star=1/(1-prob.component)
u.star
ell.min=function(w,u){min(w/u)} #already defined in preceding exercice
d=5
n_0=nrow(w.H)
#estimate the probability of the joint exceedance event:
d/n_0*sum(apply(w.H,1,ell.min,u=u.star))
#compare to observed empirical probability:
mean(apply(t(t(data.stand)/u.star)>1,1,all))
```

In this case, we see that the estimate is relatively close to the empirically observed value. Note that the estimate based on the spectral measure uses more data for the estimation and has therefore much less estimation variance than the direct empirical estimate, which can be very variable depending on the dataset.

### 3.3) Estimation of a parametric Dirichlet model for the angular measure

We first define the log-likelihood function, and then use it within the maxLik function to compute the maximum likelihood estimate.

```{r load packages 3, message=FALSE}
library(MCMCpack) #for Dirichlet density 
library(maxLik) #for numerical maximisation of the log-likelihood
```

```{r, echo=TRUE}
#define the log-likelihood function:
llfun=function(alpha){
if(alpha<=0) return(-Inf) #inadmissible parameters
 sum(log(ddirichlet(w.H,alpha=rep(alpha,d))))
}

initpar=1 # define the initial value of the parameter
fit=maxLik(llfun,start=initpar)
summary(fit)
```

Here the estimated Dirichlet shape parameter is 0.74.

### 3.4) Pairwise estimations of the  Huesler-Reiss dependence parameters

We construct a dooble loop to run through all combinations of two different variables and estimate the corresponding Huesler-Reiss distributions.

First, we load the required R packages. 
```{r load packages 4, message=FALSE}
library(evd)
```
```{r, echo=TRUE}
u=quantile(mydata,0.95)
Lambda=matrix(0,d,d)
for(j1 in 1:(d-1)){
  for(j2 in (j1+1):d){
    fit.bv=fbvpot(mydata[,c(j1,j2)],threshold=c(u,u),model="hr")
    Lambda[j1,j2]=Lambda[j2,j1]=1/fit.bv$estimate[5]
  }
}
Lambda
```

Some differences arise in the estimated values. We estimate the strongest extremal dependence between components 1 and 2, and the weakest extremal dependence between components 1 and 5. We could also check the standard errors of these estimates to see if these differences are significant.

### 2022-03-16 Practicals/Day1_MEUSE.Rmd
Saved changes to: Code_examples/plot_wind.R Practicals/Day1_MEUSE.Rmd Practicals/Day1_MEUSE.html

@@
 # import meuse data
-source("meuse.data.R")
+source("Day1_meuse.data.R")
 attach(meuse)
@@
 names(meuse)
-xyplot(Discharge~Year,pch=19)
+xyplot(Discharge~Year,pch=5)
 ```
@@
 qqmath(Discharge,xlab="Normal quantiles")
+#abline(a = 1, b = 0)
 # QQplot for Extreme Value Analysis
@@
 ```{r}
-gev.profxi(meuse.GEV)   #-.3,.3
-gev.prof(meuse.GEV,100) # 2500,4500
+gev.profxi(meuse.GEV,-.3,.3)
+gev.prof(meuse.GEV,100, 2500,4500)
 ```
@@
 identify(meuse)
+which.max(meuse$Discharge)
 ```
@@
 ```{r}
-Year[15]; Year[85]
+Year[15]; Year[83]
 
@@
 ```{r}
-mm2 = data.frame(cbind(Year[c(-15,-85)],Discharge[c(-15,-85)]))
+mm2 = data.frame(cbind(Year[c(-15,-83)],Discharge[c(-15,-83)]))
 colnames(mm2) = c("Year","Discharge")

### 2022-03-16 Code_examples/plot_wind.R

@@ -1,6 +1,8 @@
-load("team_number.RData")
+load("Datasets/1_Wind.RData")
+library(ggmap)
 
 bbox <- c(-21.375,  40.875,  16.875,  61.125)
+
 map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")
 
 dataToPlot = data.frame(X = coordinates[, 1], Y = coordinates[, 2])

### 2022-03-16 Practicals/practicalSpatialExtremes/Show/showv2.tex
Saved changes to: .gitattributes Practicals/practicalSpatialExtremes/Code/AsAnAside.R 	Practicals/practicalSpatialExtremes/Code/FitNonSimpleMaxStable.R 	Practicals/practicalSpatialExtremes/Code/SpatialTrends.R 	Practicals/practicalSpatialExtremes/Code/concprob.R 	Practicals/practicalSpatialExtremes/Code/finalFit.RData 	Practicals/practicalSpatialExtremes/Code/finalFitNew.RData 	Practicals/practicalSpatialExtremes/Code/fmadogram.R 	Practicals/practicalSpatialExtremes/Code/leastSquares.R 	Practicals/practicalSpatialExtremes/Code/modelSelection.R 	Practicals/practicalSpatialExtremes/Code/oneSimFi

\documentclass[mode=present,style=simple,paper=screen]{powerdot}

\usepackage[utf8]{inputenc}
\usepackage[square]{natbib}
\usepackage{bibentry}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{bbding}
\usepackage{manfnt}
\usepackage{fourier}
\usepackage[labelfont=bf,textfont=sl,tableposition=top,scriptsize]{caption}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{mathrsfs}

\theoremstyle{plain}% default
\newtheorem{thm}{Theorem}%[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop} {Proposition}
\newtheorem{cor}{Corollary}
\theoremstyle{definition}
\newtheorem{defn}{Definition}%[section]
\newtheorem{conj}{Conjecture}%[section]
\newtheorem{exmp}{Example}%[section]
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem{case}{Case}

\newcommand{\advice}[1]{\HandPencilLeft{} \emph{#1}}
\newcommand{\reminder}[1]{%
  \fcolorbox{blue}{white}{
    \begin{minipage}[l]{0.2\linewidth}
      \includegraphics[width=1\textwidth]{Figures/reminder}
    \end{minipage}
    \begin{minipage}[r]{0.8\linewidth}
      #1
    \end{minipage}
  }
}

\title{Statistical modelling of spatial extremes using the SpatialExtremes package}
\author{M. Ribatet}
\date{University of Montpellier}
% \date{
%   \begin{center}
%     $^\dag$ Laboratoire de Mathématiques et Application, Université de
%     Poitiers\\
%     $^\ddag$ Institut de Mathématiques et de Modélisation, Université
%     Montpellier 2
%   \end{center}
% }

\pdsetup{
  lf={The \texttt{SpatialExtremes} package},
  rf={Mathieu Ribatet},
  logohook=tr,
  logopos={.99\slidewidth,.99\slideheight},
  logocmd={},
  counters={thm,lem,prop,cor,defn,conj,exmp,case,figure},
}

\begin{document}

\maketitle

\begin{wideslide}[toc=]{Rationale for the \texttt{SpatialExtremes} package}
  \begin{quotation}
    ``The aim of the \texttt{SpatialExtremes} package is to provide
    tools for the areal modelling of extreme events. The modelling
    strategies heavily rely on the \textcolor{blue}{extreme value
      theory} and in particular \textcolor{blue}{block maxima}
    techniques---unless explicitly stated.''
  \end{quotation}

  As a consequence, most often
  \begin{itemize}
  \item the data used by the package \textcolor{blue}{have to be
      extreme}---do not pass daily values for instance;
  \item \textcolor{blue}{the marginal distribution family is fixed},
    i.e., the generalized extreme value distribution family, but you
    have hands on how within this family parameters change in space;
  \item \textcolor{blue}{the process family is fixed}, i.e.,
    max-stable processes, but you have hands on which type of
    max-stable processes to use.
  \end{itemize}
\end{wideslide}

\section[toc=0. Max-stable process]{0. About the inner structure of max-stable processes}

\begin{slide}{Spectral characterization}
  \vspace*{-1.75em}
  \onslide*{1}{
    \begin{equation*}
      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
    \end{equation*}
    where $\{\zeta_i\colon i \geq 1\}$ is a Poisson point process on
    $(0, \infty)$ with intensity measure $\mbox{d$\Lambda$}(\zeta) =
    \zeta^{-2} \mbox{d$\zeta$}$ and $Y_i$ independent copies of a
    (non-negative) stochastic process such that $\mathbb{E}\{Y(s)_+\}= 1$
    for all $s \in \mathcal{X}$.
  }%
  \onslide*{2-}{
    \begin{equation*}
      Z(s) = \max_{\varphi \in \Phi} \varphi(s), \qquad s \in \mathcal{X},
    \end{equation*}
    where $\Phi = \{\varphi_i\colon i \geq 1\}$ is a Poisson point
    process on $\mathbb{C}_0$ with an appropriate intensity measure.\vspace*{-0.5em}
    \begin{figure}
      \centering
      \includegraphics[width=0.95\textwidth]{Figures/IllustMaxStable}
    \end{figure}
  }
\end{slide}

\begin{wideslide}{Extremal functions}
  \vspace*{-2em}
  \begin{figure}
      \centering
      \onslide*{1}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-1}}%
      \onslide*{2}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-2}}%
      \onslide*{3}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-3}}%
      \onslide*{4-}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-4}}%
    \end{figure}
    \vspace*{-2.75em}
    \begin{itemize}
    \item<3-> Hidden are the random functions
      $\Phi^+ = \{\varphi_1^+, \varphi_2^+, \ldots, \varphi_k^+\}$ of
      $\Phi$ such that
      \begin{equation*}
        \varphi_j^+(s_j) = Z(s_j), \qquad j = 1, \ldots, k, \qquad
        \text{\textcolor{red}{(extremal functions)}},
      \end{equation*}\vspace*{-1.5em}
    \item<4-> and the random functions $\varphi^- \in \Phi \setminus \Phi^+$,
      i.e., satisfying
      \begin{equation*}
        \varphi^-(s_j) < Z(s_j), \qquad j = 1, \ldots, k, \qquad
        \text{\textcolor{red}{(sub-extremal functions)}}
      \end{equation*}
    \end{itemize}
\end{wideslide}

\section{1. Data and descriptive analysis}

\begin{slide}[toc=Data format,method=direct]{Required data}
  Before introducing more advanced stuff, let's talk about data
  format. It is pretty simple: 
  \begin{description}
  \item[Observations] A numeric matrix such that \textcolor{blue}{each
      row is one realization of the spatial field}---or if you prefer
    one column per site;
  \item[Coordinates] A numeric matrix such that \textcolor{blue}{each
      row is the coordinates of one site}---or if you prefer the first
    column is for instance the longitude of all sites, the second one
    latitude, \ldots
  \end{description}
{\tiny
  \begin{minipage}[l]{.49\linewidth}
\begin{verbatim}
> data
       Valkenburg  Ijmuiden  De Kooy  ...
 1971         278        NA      360  ...
 1972         334        NA      376  ...
 1973         376        NA      365  ...
 1974         314        NA      304  ...
 1975         278        NA      278  ...
 1976         350        NA      345  ...
 1977         324        NA      298  ...
 1978         298        NA      329  ...
 1979         252        NA      298  ...
 ...
\end{verbatim}%
  \end{minipage}%
  \begin{minipage}[r]{.49\linewidth}
\begin{verbatim}
> coord
                        lon        lat
 Valkenburg           4.419     52.165
 Ijmuiden             4.575     52.463
 De Kooy              4.785     52.924
 Schiphol             4.774     52.301
 Vlieland             4.942     53.255
 Berkhout             4.979     52.644
 Hoorn                5.346     53.393
 De Bilt              5.177     52.101
 ...
\end{verbatim}
  \end{minipage}
}
\end{slide}

\begin{slide}[toc=,method=direct]{Additional covariates}
  In addition to the storage of observations and coordinates, you
  might want to use additional covariates. The latter can be of two
  types:
  \begin{description}
  \item[Spatial] A numeric matrix such that \textcolor{blue}{each
      column corresponds to one spatial covariate} such as elevation,
    urban/rural, \ldots
  \item[Temporal] A numeric matrix such that \textcolor{blue}{each
      column corresponds to one temporal covariate} such as time,
    annual mean temperature, \ldots
  \end{description}
{\tiny
  \begin{minipage}[l]{.49\linewidth}
\begin{verbatim}
> spat.cov
                     alt
 Valkenburg         -0.2
 Ijmuiden            4.4
 De Kooy             0.5
 Schiphol           -4.4
 Vlieland            0.9
 Berkhout           -2.5
 Hoorn               0.5
 De Bilt             2.0
 ...
\end{verbatim}
  \end{minipage}%
  \begin{minipage}[r]{.49\linewidth}
\begin{verbatim}
> temp.cov
               nao
 1971         1.87
 1972         1.57
 1973        -0.20
 1974        -0.95
 1975        -0.46
 1976         2.34
 1977        -0.49
 1978         0.70
 1979         1.11
 ...
\end{verbatim}
  \end{minipage}
}

\advice{It is always a good idea to name your columns and rows.}
\end{slide}

\begin{slide}[toc=First look]{Inspecting data}
  \begin{itemize}
  \item As usual, you should first \textcolor{blue}{scrutinize your
      data} (weird values, encoding of missing values, check out
    factors, \ldots). But you always do that, don't you?
  \item We focus on extremes, so you may wonder
    \begin{itemize}
    \item are my data \textcolor{blue}{extremes}, i.e., block maxima?
    \item is my \textcolor{blue}{block size} appropriate?
    \item what about \textcolor{blue}{seasonality}? Refine the block
      or use a temporal covariate?
    \end{itemize}
  \item You might want to \textcolor{blue}{check that the generalized
      extreme value family is sensible for your data}---the
    \texttt{evd} package + a few lines of code will do the job for you
    (homework) \ldots 
  \item This will generally be OK, but now you have to go a bit
    further by analyzing
    \begin{itemize}
    \item the spatial dependence;
    \item and the presence/absence of any spatial trends.
    \end{itemize}
  \end{itemize}
\end{slide}

\begin{slide}[toc=Spatial dependence]{Spatial dependence \citep{Cooley2006}}
  \begin{itemize}
  \item Essentially you want to check if your data exhibit any
    (spatial) dependence. If not why would you bother with spatial
    models?
  \item The most convenient way to do this is through the
    \textcolor{blue}{$F$-madogram} and its connection with the
    \textcolor{blue}{extremal coefficient}:
    \begin{equation*}
      \nu_F(h) = \frac{1}{2} \mathbb{E} [| F\{Z(o)\} - F\{Z(h)\}| ],
      \qquad \theta(h) = \frac{1 + 2 \nu_F(h)}{1 - 2 \nu_F(h)}.
    \end{equation*}
  \item The \texttt{fmadogram} function will estimate (empirically)
    the pairwise extremal coefficent from the $F$--madogram.
  \end{itemize}

  \reminder{\begin{equation*}
      \theta(h) = - z \log \Pr\{Z(s) \leq z, Z(s + h) \leq z\}
    \end{equation*}
    where $1 \leq \theta(h) \leq 2$, with complete dependence iff
    $\theta(h) = 1$ and independence iff $\theta(h) = 2$.}
\end{slide}

\begin{slide}[toc=]{Spatial dependence (2) \citep{Dombry2017}}
  \begin{itemize}
  \item Another recent summary measure of the spatial dependence is
    the \textcolor{blue}{extremal concurrence probability function}
    $p\colon h \mapsto p(h) \in [0, 1]$ where
    \begin{equation*}
      p(h) \onslide*{3-}{\textcolor{red}{:= p(s, s+h)}} = \Pr\{\exists! \varphi \in \Phi\colon \varphi(s) = Z(s), \
      \varphi(s+h) = Z(s+h)\},
    \end{equation*}
    \pause i.e., there is a single extremal function at position $s$ and
    $s+h$.
  \end{itemize}
  \vfill
  \reminder{%
    \begin{align*}
      p(h) &= \mathbb{E}[\mbox{sign}\{Z(s) - \tilde Z(s)\}
             \mbox{sign}\{Z(s+h) - \tilde Z(s+h)\}]\\
      &= \text{\textcolor{blue}{Kendall's $\tau$}},
    \end{align*}
    and $0 \leq p(h) \leq 1$ with complete dependence iff $p(h) =
    1$ and independence iff $p(h) = 0$.
  }
  \onslide{3-}{\textcolor{red}{\HandRight{} NB: here we assume a stationary dependence structure!}}
\end{slide}

\begin{wideslide}[toc=]{The \texttt{fmadogram} function}
  \vspace{-1.5em}
  \begin{itemize}
  \item Run the file \texttt{fmadogram.R}. You should get the figure
    below.
  \item<2-> You can also use a binned version with \texttt{n.bins =
      300}\ldots
  \end{itemize}
  \begin{figure}
    \centering
    \onslide*{1}{\includegraphics[width=.75\textwidth]{Figures/fmadogram}}
    \onslide*{2-}{\includegraphics[width=.75\textwidth]{Figures/fmadogramBinned}}
    \caption{Use of the \texttt{fmadogram} function to assess the spatial dependence.}
    \label{fig:fmadogram}
  \end{figure}
  %%%\onslide{4}{\advice{Always think about what should be your spatial coordinates}}
\end{wideslide}

\begin{slide}[toc=]{The \texttt{concprob} function}
  \begin{itemize}
  \item Run the file \texttt{concprob.R}. You should get the figure
    below.
  \item<2-> You can also use a binned version with \texttt{n.bins =
      300}\ldots
  \end{itemize}
  \begin{figure}
    \centering
    \onslide*{1}{\includegraphics[width=.66\textwidth]{Figures/concprob}}
    \onslide*{2-}{\includegraphics[width=.66\textwidth]{Figures/concprobBinned}}
    \caption{Use of the \texttt{concprob} function to assess the spatial dependence.}
  \end{figure}
  %%%\onslide{4}{\advice{Always think about what should be your spatial coordinates}}
\end{slide}

\begin{wideslide}[toc=]{As an aside (\texttt{AsAnAside.R}) \citep{Dombry2017}}
  \vspace*{-1.75em}
  \begin{itemize}
  \item We can estimate the spatial distribution of the extremal
    concurrence probability \textcolor{blue}{w.r.t. a given weather
      station}, e.g., plotting
    \begin{equation*}
      \left\{(s, p(\text{Zurich}, s))\colon s \in \mathcal{X}\right\},
    \end{equation*}\vspace*{-1.5em}
  \item<2-> or estimate the expected area of \textcolor{blue}{concurrence cells}
    \begin{equation*}
      A(s_0) =
      \mathbb{E}\left\{\int_\mathcal{X} 1_{\{\text{$s_0$ and $s$ are
            concurrent}\}} \mbox{d$s$} \right\} = \int_\mathcal{X}
        p(s_0, s) \mbox{d$s$}
    \end{equation*}
  \end{itemize}
  \vspace*{-1.5em}
  \begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/concMap1}\hfill%
    \onslide{2}{\includegraphics[width=0.45\textwidth]{Figures/concMap2}}%
  \end{figure}
\end{wideslide}

\begin{wideslide}[toc=Spatial trends]{Spatial trends}
  %\vspace*{-1.5em}
  \begin{itemize}
  \item We can do a \textcolor{blue}{symbol plot} see the file \texttt{SpatialTrends.R}.
  \end{itemize}
  \pause
  \begin{figure}
    \centering
    \onslide*{2}{\includegraphics[width=0.5\textwidth]{Figures/symbolPlot}}
    \onslide*{3-}{\includegraphics[width=\textwidth]{Figures/symbolPlot2}}
    \caption{Symbol plot for the Swiss precipitation data.}
  \end{figure}\vfill
  \onslide*{4-}{\advice{When exporting figures into eps/pdf, always
      pay attention to the aspect ratio.}}
\end{wideslide}

\begin{slide}[toc=Debrief \#1]{What we have learned so far (apart from
    using \texttt{SpatialExtremes})}
  \begin{itemize}
  \item The data exhibit some \textcolor{blue}{spatial
      dependence} and there is still some (weak) dependance at a
    separation lag of 100km.
  \item There's a clear \textcolor{blue}{north-west / south-east
      gradient} in the intensities of rainfall storms.
  \item In conclusion it makes sense to use max-stable
    \textcolor{blue}{processes} whose \textcolor{blue}{marginal
      parameters are not constant across space}.
  \item More specifically, we have:
    \begin{itemize}
    \item a clear north-west / south-east gradient for the location
      and scale parameters;
    \item no clear pattern for the shape parameter.
    \end{itemize}
  \end{itemize}
\end{slide}

\section{2. Simple max-stable processes}

\begin{slide}{Max-stable models}
  \begin{itemize}
  \item In this section we \textcolor{blue}{focus only on the spatial
      dependence} and so assume that the margins are known and
    \textcolor{blue}{unit Fréchet}---this is a standard choice in
    extreme value theory.
  \item From the spectral characterization
    \begin{equation*}
      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
    \end{equation*}
    we can propose several parametric models for spatial
    extremes. Hence by letting $Y$ to be
    \begin{description}
    \item[Gaussian densities] with random displacements we get the
      \textcolor{blue}{Smith} process;
    \item[Gaussian] we get the \textcolor{blue}{Schlather} process;
    \item[Log-normal] (with a drift) we get the \textcolor{blue}{Brown--Resnick} process;
    \item[Gaussian] but elevated to some power we get the
      \textcolor{blue}{Extremal-$t$} process.
    \end{description}
  \end{itemize}
\end{slide}

\begin{slide}[toc=]{Dependence parameters}
  \begin{description}
  \item[Smith] Elements of the covariance matrix appearing in the
    Gaussian densities;
  \item[Schlather] Parameters of the correlation function;
  \item[Brown--Resnick] Parameters of the semi-variogram;
  \item[Extremal--$t$] Parameters of the correlation function and
    degrees of freedom.
  \end{description}
  \begin{itemize}
  \item Since the margins are fixed, we only need to get estimates for
    the dependence parameters.
  \item How can we do that?
  \end{itemize}
\end{slide}

\begin{wideslide}[toc=Least squares,method=file]{Least squares
    (\texttt{leastSquares.R}) \citep{Smith1990}}
  \vspace*{-0.5em}
  \begin{equation*}
    \underset{\psi \in \Psi}{\arg \min} \sum_{1 \leq i < j \leq k}
    \left\{\theta(s_j - s_j; \psi) - \hat \theta(s_i - s_j)
    \right\}^2,
  \end{equation*}
  where $\theta(\cdot; \psi)$ is the extremal coefficient obtained
  from the max-stable model with dependence parameters set to $\psi$
  and $\hat \theta(\cdot)$ is any empirical estimates of the extremal
  coefficient, e.g., $F$--madogram based.

  \begin{minipage}[l]{.33\linewidth}
{\tiny
\begin{verbatim}
> M0
        Estimator: Least Squares
            Model: Schlather
         Weighted: TRUE
  Objective Value: 3592.429
Covariance Family: Whittle-Matern

Estimates
  Marginal Parameters:
  Assuming unit Frechet.

  Dependence Parameters:
  range   smooth
54.3239   0.4026

Optimization Information
  Convergence: successful
  Function Evaluations: 61
\end{verbatim}
}
  \end{minipage}
  \hfill
  \begin{minipage}[r]{.6\linewidth}
    \begin{figure}
      \centering
      \includegraphics[width=0.49\textwidth]{Figures/empExtCoeffRain}%
      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/leastSquares}}
      \caption{Fitting simple max-stable processes from least
        squares.}
      \label{fig:leastSquares}
    \end{figure}
  \end{minipage}
\end{wideslide}

\begin{wideslide}[toc=Pairwise likelihood,method=file]{Pairwise
    likelihood (\texttt{pairwiseLlik.R}) \citep{Padoan2010}}
  \vspace*{-2em}
  \begin{equation*}
    \underset{\psi \in \Psi}{\arg \max} \sum_{\ell = 1}^n \sum_{1 \leq
      i < j \leq k} \log f\{z_\ell(s_i), z_\ell(s_j); \psi\},
  \end{equation*}
  where $f(\cdot, \cdot; \psi)$ is the bivariate density of the
  considered max-stable model.

    \begin{minipage}[l]{.33\linewidth}
      {\tiny
\begin{verbatim}
        Estimator: MPLE
            Model: Schlather
         Weighted: FALSE
   Pair. Deviance: 1136863
              TIC: 1137456
Covariance Family: Whittle-Matern

Estimates
  Marginal Parameters:
  Assuming unit Frechet.

  Dependence Parameters:
  range   smooth
50.1976   0.3713

Standard Errors
  range   smooth
20.7085   0.0789

Asymptotic Variance Covariance
        range       smooth
range   428.841018   -1.570081
smooth   -1.570081    0.006225
Optimization Information
  Convergence: successful
  Function Evaluations: 67
\end{verbatim}
}
  \end{minipage}
  \hfill
  \begin{minipage}[r]{.6\linewidth}
    \begin{figure}
      \centering
      \includegraphics[width=0.49\textwidth]{Figures/empExtCoeffRain}%
      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/pairwiseLlik}}
      \caption{Fitting simple max-stable processes maximizing pairwise
        likelihood.}
      \label{fig:pairLlik}
    \end{figure}
  \end{minipage}
\end{wideslide}

\begin{slide}[method=direct,toc= Model selection]{Model Selection \citep{Varin2005}}
  \begin{itemize}
  \item The advantage of the pairwise likelihood estimator over the
    least squares one is that you can do \textcolor{blue}{model
      selection}.
  \item For instance one can use the ${\rm TIC}$,
    \textcolor{blue}{Takeuchi Information Criterion} or sometimes
    known as ${\rm CLIC}$, Composite Likelihood Information Criterion,
    \begin{equation*}
      {\rm TIC} = 2 \ell_{\rm pairwise}(\hat \psi) - 2
      \mbox{tr}\{J(\hat \psi) H^{-1} (\hat \psi)\},
    \end{equation*}
    $H(\hat\psi) = \mathbb{E} \{\nabla^2 \ell_{\rm pairwise}(Y; \hat
    \psi) \}$, $J(\hat \psi) = \mbox{Var}\{\nabla \ell_{\rm
      pairwise}(Y; \hat\psi)\}$.
  \item From our previous fitted models, we get
{\tiny
\begin{verbatim}
> TIC(M0,M1,M2)
     M1      M2      M0
1133660 1134823 1137449
\end{verbatim}
}
  \end{itemize}
\end{slide}

\begin{slide}[toc=Simulation,method=direct]{Simulating simple max-stable processes
    (\texttt{simulation.R}) \citep{Schlather2002,Dombry2016}}
  \begin{itemize}
  \item Once you have fitted a suitable model, you usually want to
    simulate from it.
  \item Simulation from max-stable models is rather complex, recall
    that
    \begin{equation*}
      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X}.
    \end{equation*}
  \end{itemize}
  \begin{minipage}[l]{.6\linewidth}
{\tiny
\begin{verbatim}
sim <- rmaxstab(n.obs, cbind(x, y), "twhitmat", DoF = 4,
+               nugget = 0, range = 3, smooth = 1)
> sim
            [,1]       [,2]      [,3]      [,4]       [,5]
 [1,]  3.8048914  0.4767980 6.3613989 1.4548317  1.0433912
 [2,]  1.2200332  0.6711422 0.8078701 2.0928629  0.7537061
 [3,]  0.5466466  2.0498561 4.8852572 2.3497976  0.6857268
 [4,]  ...
\end{verbatim}
}
  \end{minipage}
  \hfill
  \begin{minipage}[r]{.33\linewidth}
    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{Figures/simulation}
      \caption{One simulation on a 50 x 50 grid from the extremal--$t$
        model. (log scale)}
    \end{figure}
  \end{minipage}
\end{slide}

\begin{slide}[toc=Debrief \#2]{What we have learned so far (apart from
    using \texttt{SpatialExtremes})}
  \begin{itemize}
  \item The Smith model is clearly not a sensible model for our
    data---because of its linear behaviour near the origin;
  \item Schlather, Brown--Resnick and Extremal--$t$ seems relevant;
  \item According to the ${\rm TIC}$, the Extremal--$t$ should be
    preferred.
  \end{itemize}
\end{slide}

\section{3. Trends surfaces}

\begin{slide}[toc=]{From generalized extreme value margins to unit
    Fréchet ones}
  \begin{itemize}
  \item Alright! We are able to handle the spatial dependence, but
    \textcolor{blue}{we assume that our data have unit Fréchet
      margins}. This is not realistic at all!
  \item Fortunately, if $Y \sim {\rm GEV}(\mu, \sigma, \xi)$ then
    \begin{equation*}
      Z = \left(1 + \xi \frac{Y - \mu}{\sigma} \right)^{1 / \xi} \sim
      \text{Unit Fréchet}=\mbox{GEV}(1,1,1).
    \end{equation*}
    \pause
  \item And since we are extreme value and spatial guys
    \begin{equation*}
      Z\textcolor{blue}{(s)} = \left\{1 + \xi\textcolor{blue}{(s)}
        \frac{Y\textcolor{blue}{(s)} -
          \mu\textcolor{blue}{(s)}}{\sigma\textcolor{blue}{(s)}}
      \right\}^{1 / \xi\textcolor{blue}{(s)}}, \qquad s \in \mathcal{X},
    \end{equation*}
    is a simple max-stable process.
  \item Hence we can use the maximum pairwise likelihood estimator as
    before---up to an \textcolor{blue}{additional Jacobian term}.
  \end{itemize}
\end{slide}

\begin{slide}[toc=Spatial GEV]{Omitting the spatial dependence}
  \begin{itemize}
  \item With simple max-stable models, we omitted the marginal parameters.
  \item Here we will \textcolor{blue}{omit the spatial dependence for
      a while} and consider locations as being
    \textcolor{blue}{mutually independent}, i.e., use independence
    likelihood
    \begin{equation*}
      \underset{\psi \in \Psi}{\arg \max} \sum_{i=1}^k \ell_{\rm
        GEV}\{y(s_i); \psi\}.
    \end{equation*}
  \item This is a kind of ``\emph{spatial GEV}'' where $\psi$ is a
    vector of marginal parameters.
  \end{itemize}
\end{slide}

\begin{slide}[toc=,method=direct]{Defining trend surfaces}
  \vspace*{-1em}
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/symbolPlot2}
    \caption{Symbol plot for the Swiss precipitation data.}
  \end{figure}
  This suggests that
  \begin{align*}
    \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
      \mu} {\rm lat}(x) + \beta_{3, \mu} {\rm lon}(s) \times {\rm lat}(s),\\
    \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
      \sigma} {\rm lat}(s) + \beta_{3, \sigma} {\rm lon}(s) \times {\rm lat}(s),\\
    \xi(s) &= \beta_{0,\xi},
  \end{align*}
  or equivalently with the \texttt{R} language\vspace*{-0.5em}
{\small
\begin{verbatim}
loc.form <- scale.form <- y ~ lon * lat; shape.form <- y ~ 1
\end{verbatim}
}
\end{slide}

\begin{slide}[toc=,method=direct]{Fitting the \emph{spatial GEV} model
    (\texttt{spatialGEV.R}) \citep{Davison2012}}
  \vspace*{-1em}
  \tiny{
\begin{verbatim}
      Model: Spatial GEV model
   Deviance: 29303.81
        TIC: 29499.38

    Location Parameters:
locCoeff1  locCoeff2  locCoeff3  locCoeff4
   27.132      1.846     -3.656     -1.080
       Scale Parameters:
scaleCoeff1  scaleCoeff2  scaleCoeff3  scaleCoeff4
     9.7850       0.7023      -1.0858      -0.5531
       Shape Parameters:
shapeCoeff1
     0.1572

Standard Errors
  locCoeff1    locCoeff2    locCoeff3    locCoeff4  scaleCoeff1  scaleCoeff2
    1.13326      0.34864      0.45216      0.38361      0.76484      0.28446
scaleCoeff3  scaleCoeff4  shapeCoeff1
    0.31267      0.27566      0.05878

Asymptotic Variance Covariance
             locCoeff1   locCoeff2   locCoeff3   locCoeff4   scaleCoeff1
locCoeff1     1.2842711   0.1131400  -0.1740921  -0.0729564   0.6570988
locCoeff2     0.1131400   0.1215498  -0.0623759   0.0149596   0.0521630
locCoeff3    -0.1740921  -0.0623759   0.2044448   0.0576622  -0.1086629
locCoeff4    -0.0729564   0.0149596   0.0576622   0.1471593  -0.0346376
scaleCoeff1   0.6570988   0.0521630  -0.1086629  -0.0346376   0.5849729
...

Optimization Information
  Convergence: successful
  Function Evaluations: 2135
\end{verbatim}
    }
\end{slide}

\begin{wideslide}[toc=Prediction \#1]{Get predictions (\texttt{predictionSpatialGEV.R})}
  \begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Figures/symbolPlot}%
    \includegraphics[width=.5\textwidth]{Figures/predictionSpatialGEV}
    \caption{Left: symbol plot. Right: Prediction of the pointwise
      25-year return levels from a fitted spatial GEV model.}
    \label{fig:prediction}
  \end{figure}
  \pause
  \begin{itemize}
  \item But don't we forget something??? \pause
  \item Model selection?
  \end{itemize}
\end{wideslide}

\begin{slide}[toc=Model selection \#2,method=direct]{Model selection \#2
    (\texttt{modelSelection.R}) \citep{Chandler2007,Kent1982}}
  \begin{itemize}
  \item Typically here we would like to test if a given covariate is
    required or not
  \item Hence we're dealing with nested model for which
    \textcolor{blue}{composite likelihood ratio test} are especially
    suited
    \begin{equation*}
      2 \{\ell_{\rm composite}(\hat \psi) - \ell_{\rm composite}(\hat
      \phi_{\lambda_0}, \lambda_0)\} \longrightarrow \sum_{j=1}^p
      \lambda_j X_i, \qquad n \to \infty.
    \end{equation*}
  \end{itemize}
{\tiny
\begin{verbatim}
Eigenvalue(s): 2.7 1.95

Analysis of Variance Table
   MDf Deviance Df  Chisq Pr(> sum lambda Chisq)
M2   7    29328
M0   9    29306  2 22.265               0.008273 **
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
\end{verbatim}
}
\advice{Always check that your models are nested. The code won't do
  that for you!}
\end{slide}

\begin{slide}[toc=Debrief \#3]{What we have learned so far (apart from
    using \texttt{SpatialExtremes})}
  \begin{itemize}
  \item Based on the \texttt{spatial GEV} model, we identify what
    seems to be relevant trend surfaces for the marginal parameters:
    \begin{align*}
      \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
        \mu} {\rm lat}(s) + \beta_{3,\mu} {\rm lon}(s) {\rm lat(s)},\\
      \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
        \sigma} {\rm lat}(s),\\
      \xi(s) &= \beta_{0,\xi},
    \end{align*}
  \end{itemize}
\end{slide}

\section{4. General max-stable processes}

\begin{wideslide}[toc=Fitting,method=direct]{Fitting a max-stable process with trend
    surfaces}
  \begin{itemize}
  \item Now it's time to combine everything, i.e.,
    \textcolor{blue}{trend surfaces + dependence}.
  \item The syntax won't be a big surprise
  \end{itemize}
  {\tiny
\begin{verbatim}
M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form, scale.form, shape.form)

       Estimator: MPLE
            Model: Extremal-t
         Weighted: FALSE
   Pair. Deviance: 2237562
              TIC: 2249206
Covariance Family: Whittle-Matern

Estimates
  Marginal Parameters:
    Location Parameters:
locCoeff1  locCoeff2  locCoeff3  locCoeff4
27.136295   0.060145  -0.164755  -0.001117
       Scale Parameters:
scaleCoeff1  scaleCoeff2  scaleCoeff3
    9.88857      0.02869     -0.04581
       Shape Parameters:
shapeCoeff1
     0.1727
  Dependence Parameters:
   range    smooth       DoF
225.9452    0.3645    4.1566
...
\end{verbatim}
}
\end{wideslide}

\begin{wideslide}[method=direct]{Model checking \citep{Davison2012}}
  \begin{minipage}[l]{.49\linewidth}
    \begin{itemize}
    \item When you want to check your fitted max-stable model, you
      usually want to check if
      \begin{itemize}
      \item observations at each single location are well modelled:
        \textcolor{blue}{return level plot};
      \item the dependence is well captured: \textcolor{blue}{extremal
          coefficient function}.
      \end{itemize}
    \item This can be done using a single line of code
\begin{verbatim}
> plot(M0)
\end{verbatim}
    \end{itemize}
  \end{minipage}
  \hfill
  \onslide{2}{
  \begin{minipage}[r]{.49\linewidth}
    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{Figures/checkFinalModel}
      \caption{Model checking for a fitted max-stable process having
        trend surfaces.}
    \end{figure}
  \end{minipage}
}
\end{wideslide}

\begin{wideslide}{Predictions (\texttt{simulationFinal.R})}
  \begin{itemize}
  \item Prediction works as for the \emph{spatial GEV model} thanks to
    the \texttt{predict} function.
  \item But beware these predictions are pointwise---no spatial
    dependence at all!!!
  \item If you want to do take into account spatial dependence then
    you need to simulate from your fitted model.
  \end{itemize}
  \vspace*{-2em}
  \begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{Figures/oneSimFinalModel}
    \caption{One simulation from our fitted extremal--$t$ model with trend surfaces.}
  \end{figure}
\end{wideslide}

\section{5. Conclusion}

\begin{wideslide}[toc=]{What we haven't seen}
  \begin{itemize}
  \item Using weighted pairwise likelihood;
  \item Many (many!) utility functions. Highly recommended to have a
    look at the documentation;
  \item The package has a vignette: \texttt{vignette("SpatialExtremesGuide")};
  \item Copula models---although I do not recommend their use for
    spatial extremes;
  \item Bayesian hierarchical models;
  \item Unconditional simulations: several implementations (including
    exact simulations)
  \item Conditional simulations---really CPU demanding.
  \end{itemize}
  \pause
  \HandRight{} A rather recent review on max-stable processes with R
  code is given by \citet{Ribatet2013b}
  \vfill
  \begin{center}
    \Large THANK YOU!
  \end{center}
  \vfill
\end{wideslide}

\begin{wideslide}{References}
  \vspace*{-1.5em}
  {\footnotesize
  \bibliographystyle{apalike}
  \nobibliography{/Users/Mathieu/Dropbox/biblio_ribatet}

  \begin{itemize}
  \item[\checkmark] \bibentry{Chandler2007}
  \item[\checkmark] \bibentry{Cooley2006}
  \item[\checkmark] \bibentry{Davison2012}
  \item[\checkmark] \bibentry{Dombry2016}
  \item[\checkmark] \bibentry{Dombry2017}
  \item[\checkmark] \bibentry{Padoan2010}
  \item[\checkmark] \bibentry{Ribatet2013b}
  \item[\checkmark] \bibentry{Schlather2002}
  \item[\checkmark] \bibentry{Smith1990}
  \item[\checkmark] \bibentry{Varin2005}
  \end{itemize}
  }
\end{wideslide}

%\bibliography{/Users/Mathieu/Dropbox/biblio_ribatet}
%\bibliographystyle{apalike}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/FitNonSimpleMaxStable.R

@@ -1,0 +1,16 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Define the trend surfaces
+loc.form <- y ~ lon * lat
+scale.form <- y ~ lon + lat
+shape.form <- y ~ 1
+
+## Center and scale the covariates
+coord <- scale(coord[,1:2], scale = FALSE)
+M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form,
+                 scale.form, shape.form)
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/concprob.R

@@ -1,0 +1,17 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## These data are maximum daily rainfall amounts over the years
+## 1962–2008 occuring during June–August at 79 sites in Switzerland.
+
+ls() ## <<-- we have two R objects 'coord' (the spatial coordinates)
+     ##      and 'rain' (annual maxima precipitation)
+
+## Use the concprob function to assess the spatial dependence (if any)
+concprob(rain, coord[,1:2])
+concprob(rain, coord[,1:2], n.bins = 300)
+
+## Please feel free to give me your comments !!!

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/fmadogram.R

@@ -1,0 +1,17 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## These data are maximum daily rainfall amounts over the years
+## 1962–2008 occuring during June–August at 79 sites in Switzerland.
+
+ls() ## <<-- we have two R objects 'coord' (the spatial coordinates)
+     ##      and 'rain' (annual maxima precipitation)
+
+## Use the fmadogram function to assess the spatial dependence (if any)
+fmadogram(rain, coord[,1:2])
+fmadogram(rain, coord[,1:2], n.bins = 300)
+
+## Please feel free to give me your comments !!!

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/leastSquares.R

@@ -1,0 +1,21 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Fit a bunch of max-stable models using least squares
+M0 <- lsmaxstab(rain, coord[,1:2], "whitmat", nugget = 0) ##<<-- Schlather model with Whittle-Matern corr. fun.
+M1 <- lsmaxstab(rain, coord[,1:2], "gauss", iso = TRUE)  ##<<-- Isotropic Smith model
+M2 <- lsmaxstab(rain, coord[,1:2], "tpowexp", nugget = 0) ## <<-- Extremal-t with powered exponential corr. fun.
+M3 <- lsmaxstab(rain, coord[,1:2], "brown") ## <<-- Brown--Resnick process
+
+## Compare the empirical extremal coefficient cloud with fitted extremal coefficient function
+fmadogram(rain, coord[,1:2], which = "ext", col = "grey")
+fmadogram(fitted = M0, which = "ext", n.bins = 200, add = TRUE)## <<- you can pass a fitted model also
+plot(M1$ext.coeff, from = 0, to = 120, col = 3, add = TRUE)
+plot(M2$ext.coeff, from = 0, to = 120, col = 4, add = TRUE)
+plot(M3$ext.coeff, from = 0, to = 120, col = 5, add = TRUE)
+legend("bottomright", c("Schlather", "Smith", "Extremal-t", "Brown-Resnick"),
+       col = 2:5, lty = 1, bty = "n", inset = 0.05)
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/modelSelection.R

@@ -1,0 +1,23 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Define the trend surfaces
+loc.form <- scale.form <- y ~ lon * lat
+shape.form <- y ~ 1
+
+## Center and scale the covariates
+covariates <- scale(coord[,1:2], scale = FALSE)
+M0 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form)
+
+## Consider nested models, i.e., no lon / lat interaction
+M1 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form,
+                 scaleCoeff4 = 0)
+M2 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form,
+                 locCoeff4 = 0, scaleCoeff4 = 0)
+
+anova(M0, M1)## <<-- p-value really small so what can you conclude???
+anova(M1, M2)## <<-- the same here??? So what?
+anova(M0, M2)## <<-- Is this coherent from the above two tests?

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/predictionSpatialGEV.R

@@ -1,0 +1,41 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Define the trend surfaces
+loc.form <- scale.form <- y ~ lon * lat
+shape.form <- y ~ 1
+
+## Center and scale the covariates
+covariates <- scale(coord[,1:2], scale = FALSE)
+M0 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form)
+
+## Define the grid where we want prediction
+n.grid <- 250
+lon.grid <- seq(645, 770, length = n.grid)
+lat.grid <- seq(209, 291, length = n.grid)
+
+grid <- expand.grid(lon.grid, lat.grid)
+colnames(grid) <- c("lon", "lat") ## <<-- I told you it's important to
+                                  ##      name columns and rows ;-)
+
+## Since we scale the covariates we need to do the same with our grid
+grid.scaled <- grid
+for (i in 1:2)
+    grid.scaled[,i] <- grid.scaled[,i] - attributes(covariates)$"scaled:center"[i]
+
+## Get the prediction for the 25-year return level
+predictions <- predict(M0, grid.scaled, ret.per = 25)
+
+## Plot it (yes I know I should write a function that do that
+## automagically)
+col.palette <- colorRampPalette(c("blue", "white", "red"))
+
+swiss()
+image(lon.grid, lat.grid, matrix(predictions[,"Q25"], n.grid), col = col.palette(64),
+      add = TRUE)
+contour(lon.grid, lat.grid, matrix(predictions[,"Q25"], n.grid), add = TRUE)
+swiss(add = TRUE)
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/simulation.R

@@ -1,0 +1,22 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## Simulate from an extremal-t model at non gridded data
+n.site <- 5
+x <- runif(n.site, 0, 10)
+y <- runif(n.site, 0, 10)
+
+n.obs <- 20
+sim <- rmaxstab(n.obs, cbind(x, y), "twhitmat", DoF = 4,
+                nugget = 0, range = 3, smooth = 1)
+
+
+## Simulate from an extremal-t model at gridded data
+n.site <- 50
+x <- y <- seq(0, 10, length = n.site)
+
+sim2 <- rmaxstab(1, cbind(x, y), "twhitmat", DoF = 4,
+                nugget = 0, range = 3, smooth = 1, grid = TRUE)
+
+## Plot the realization with a log-scale for visualization purposes
+filled.contour(log(sim2))

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/AsAnAside.R
Saved changes to: Practicals/practicalSpatialExtremes/Code/AsAnAside.R 	Practicals/practicalSpatialExtremes/Code/FitNonSimpleMaxStable.R 	Practicals/practicalSpatialExtremes/Code/SpatialTrends.R 	Practicals/practicalSpatialExtremes/Code/concprob.R 	Practicals/practicalSpatialExtremes/Code/finalFit.RData 	Practicals/practicalSpatialExtremes/Code/finalFitNew.RData 	Practicals/practicalSpatialExtremes/Code/fmadogram.R 	Practicals/practicalSpatialExtremes/Code/leastSquares.R 	Practicals/practicalSpatialExtremes/Code/modelSelection.R 	Practicals/practicalSpatialExtremes/Code/oneSimFinalModel.RData

@@ -1,13 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Estimate concurrence probabilities w.r.t. Zurich
-idx.station <- 78##78th station corresponds to Zurich
-concurrencemap(rain, coord[,1:2], type = idx.station,
-               plot.border = function(add) swiss(city = TRUE, add = add))
-
-## Estimate the expected area of pointwise concurrence cells (default)
-concurrencemap(rain, coord[,1:2], plot.border = swiss)

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/FitNonSimpleMaxStable.R

@@ -1,16 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Define the trend surfaces
-loc.form <- y ~ lon * lat
-scale.form <- y ~ lon + lat
-shape.form <- y ~ 1
-
-## Center and scale the covariates
-coord <- scale(coord[,1:2], scale = FALSE)
-M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form,
-                 scale.form, shape.form)
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/SpatialTrends.R

@@ -1,17 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-data(rainfall)
-
-## We are going to plot a bunch of circle centered at weather
-## stations, whose radius is proportional to the deviation from the
-## areal mean and whose color is blue or red depending if we're below
-## or above the areal mean
-
-## Symbol plot based on pointwise medians
-symbolplot(rain, coord[,1:2], which = "median", plot.border = swiss,
-           scale = 5, plot.legend = FALSE)
-
-## Symbol plot based on the pointwise GEV estimates (default)
-symbolplot(rain, coord[,1:2], plot.border = swiss,
-           scale = 5, plot.legend = FALSE)
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/concprob.R

@@ -1,17 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## These data are maximum daily rainfall amounts over the years
-## 1962–2008 occuring during June–August at 79 sites in Switzerland.
-
-ls() ## <<-- we have two R objects 'coord' (the spatial coordinates)
-     ##      and 'rain' (annual maxima precipitation)
-
-## Use the concprob function to assess the spatial dependence (if any)
-concprob(rain, coord[,1:2])
-concprob(rain, coord[,1:2], n.bins = 300)
-
-## Please feel free to give me your comments !!!

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/fmadogram.R

@@ -1,17 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## These data are maximum daily rainfall amounts over the years
-## 1962–2008 occuring during June–August at 79 sites in Switzerland.
-
-ls() ## <<-- we have two R objects 'coord' (the spatial coordinates)
-     ##      and 'rain' (annual maxima precipitation)
-
-## Use the fmadogram function to assess the spatial dependence (if any)
-fmadogram(rain, coord[,1:2])
-fmadogram(rain, coord[,1:2], n.bins = 300)
-
-## Please feel free to give me your comments !!!

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/leastSquares.R

@@ -1,21 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Fit a bunch of max-stable models using least squares
-M0 <- lsmaxstab(rain, coord[,1:2], "whitmat", nugget = 0) ##<<-- Schlather model with Whittle-Matern corr. fun.
-M1 <- lsmaxstab(rain, coord[,1:2], "gauss", iso = TRUE)  ##<<-- Isotropic Smith model
-M2 <- lsmaxstab(rain, coord[,1:2], "tpowexp", nugget = 0) ## <<-- Extremal-t with powered exponential corr. fun.
-M3 <- lsmaxstab(rain, coord[,1:2], "brown") ## <<-- Brown--Resnick process
-
-## Compare the empirical extremal coefficient cloud with fitted extremal coefficient function
-fmadogram(rain, coord[,1:2], which = "ext", col = "grey")
-fmadogram(fitted = M0, which = "ext", n.bins = 200, add = TRUE)## <<- you can pass a fitted model also
-plot(M1$ext.coeff, from = 0, to = 120, col = 3, add = TRUE)
-plot(M2$ext.coeff, from = 0, to = 120, col = 4, add = TRUE)
-plot(M3$ext.coeff, from = 0, to = 120, col = 5, add = TRUE)
-legend("bottomright", c("Schlather", "Smith", "Extremal-t", "Brown-Resnick"),
-       col = 2:5, lty = 1, bty = "n", inset = 0.05)
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/modelSelection.R

@@ -1,23 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Define the trend surfaces
-loc.form <- scale.form <- y ~ lon * lat
-shape.form <- y ~ 1
-
-## Center and scale the covariates
-covariates <- scale(coord[,1:2], scale = FALSE)
-M0 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form)
-
-## Consider nested models, i.e., no lon / lat interaction
-M1 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form,
-                 scaleCoeff4 = 0)
-M2 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form,
-                 locCoeff4 = 0, scaleCoeff4 = 0)
-
-anova(M0, M1)## <<-- p-value really small so what can you conclude???
-anova(M1, M2)## <<-- the same here??? So what?
-anova(M0, M2)## <<-- Is this coherent from the above two tests?

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/pairwiseLlik.R

@@ -1,25 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Transform the data to unit Fréchet first
-frechet <- apply(rain, 2, gev2frech, emp = TRUE) ## <<-- we use the empirical CDF
-
-## Fit a bunch of max-stable models using pairwise likelihood
-M0 <- fitmaxstab(frechet, coord[,1:2], "whitmat", nugget = 0)
-M1 <- fitmaxstab(frechet, coord[,1:2], "tpowexp", nugget = 0)
-M2 <- fitmaxstab(frechet, coord[,1:2], "brown")
-
-## Compare the empirical extremal coefficient cloud with fitted extremal coefficient function
-par(mar = c(4, 5, 0.5, 0.5), ps = 14)
-concprob(rain, coord[,1:2], col = "grey")
-concprob(fitted = M0, n.bins = 200, add = TRUE)## <<- you can pass a fitted model also
-plot(M1$conc.prob, from = 0, to = 120, col = 3, add = TRUE)
-plot(M2$conc.prob, from = 0, to = 120, col = 4, add = TRUE)
-legend("topright", c("Schlather", "Extremal-t", "Brown-Resnick"),
-       col = 2:4, lty = 1, bty = "n", inset = 0.05)
-
-
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/predictionSpatialGEV.R

@@ -1,41 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Define the trend surfaces
-loc.form <- scale.form <- y ~ lon * lat
-shape.form <- y ~ 1
-
-## Center and scale the covariates
-covariates <- scale(coord[,1:2], scale = FALSE)
-M0 <- fitspatgev(rain, covariates, loc.form, scale.form, shape.form)
-
-## Define the grid where we want prediction
-n.grid <- 250
-lon.grid <- seq(645, 770, length = n.grid)
-lat.grid <- seq(209, 291, length = n.grid)
-
-grid <- expand.grid(lon.grid, lat.grid)
-colnames(grid) <- c("lon", "lat") ## <<-- I told you it's important to
-                                  ##      name columns and rows ;-)
-
-## Since we scale the covariates we need to do the same with our grid
-grid.scaled <- grid
-for (i in 1:2)
-    grid.scaled[,i] <- grid.scaled[,i] - attributes(covariates)$"scaled:center"[i]
-
-## Get the prediction for the 25-year return level
-predictions <- predict(M0, grid.scaled, ret.per = 25)
-
-## Plot it (yes I know I should write a function that do that
-## automagically)
-col.palette <- colorRampPalette(c("blue", "white", "red"))
-
-swiss()
-image(lon.grid, lat.grid, matrix(predictions[,"Q25"], n.grid), col = col.palette(64),
-      add = TRUE)
-contour(lon.grid, lat.grid, matrix(predictions[,"Q25"], n.grid), add = TRUE)
-swiss(add = TRUE)
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/simulation.R

@@ -1,22 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## Simulate from an extremal-t model at non gridded data
-n.site <- 5
-x <- runif(n.site, 0, 10)
-y <- runif(n.site, 0, 10)
-
-n.obs <- 20
-sim <- rmaxstab(n.obs, cbind(x, y), "twhitmat", DoF = 4,
-                nugget = 0, range = 3, smooth = 1)
-
-
-## Simulate from an extremal-t model at gridded data
-n.site <- 50
-x <- y <- seq(0, 10, length = n.site)
-
-sim2 <- rmaxstab(1, cbind(x, y), "twhitmat", DoF = 4,
-                nugget = 0, range = 3, smooth = 1, grid = TRUE)
-
-## Plot the realization with a log-scale for visualization purposes
-filled.contour(log(sim2))

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/simulationFinal.R

@@ -1,69 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Define the trend surfaces
-loc.form <- y ~ lon.scaled * lat.scaled
-scale.form <- y ~ lon.scaled + lat.scaled
-shape.form <- y ~ 1
-
-## Center and scale the covariates
-coord.centered <- scale(coord[,1:2], scale = FALSE)
-
-## You can fit the model by invoking the following code
-##
-##M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form,
-##                 scale.form, shape.form, marg.cov = covariates)
-##
-## but for time reasons just do
-load("finalFit.RData")
-
-## Define the grid where we want prediction
-n.grid <- 50
-lon.grid <- seq(645, 770, length = n.grid)
-lat.grid <- seq(209, 291, length = n.grid)
-
-grid <- expand.grid(lon.grid, lat.grid)
-colnames(grid) <- c("lon", "lat") ## <<-- I told you it's important to
-                                  ##      name columns and rows ;-)
-
-## Since we scale the covariates we need to do the same with our grid
-grid.scaled <- grid
-for (i in 1:2)
-    grid.scaled[,i] <- (grid.scaled[,i] - attributes(coord)$"scaled:center"[i])
-
-## Get the prediction for the GEV parameters
-predictions <- predict(M0, grid.scaled)
-loc.pred <- matrix(predictions[,"loc"], n.grid)
-scale.pred <- matrix(predictions[,"scale"], n.grid)
-shape.pred <- matrix(predictions[,"shape"], n.grid)
-
-## Simulate from our fitted model
-
-## First with unit Frechet margins
-
-## For time reasons again...
-##sim <- rmaxstab(1, cbind(lon.grid, lat.grid), "twhitmat",
-##                DoF = M0$param["DoF"], nugget = M0$param["nugget"],
-##                range = M0$param["range"], smooth = M0$param["smooth"],
-##                grid = TRUE)
-load("oneSimFinalModel.RData")
-
-
-## Switch to the appropriate GEV margins
-for (i in 1:n.grid)
-    for (j in 1:n.grid)
-        sim[i,j] <- frech2gev(sim[i,j], loc.pred[i,j], scale.pred[i,j], shape.pred[i,j])
-
-## Plot the realization
[…]

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/spatialGEV.R

@@ -1,22 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## First load some data (already formatted)
-data(rainfall)
-
-## Define the trend surfaces
-loc.form <- scale.form <- y ~ lon * lat
-shape.form <- y ~ 1
-
-## Rmk: In the formula above we use lon and lat since these are the
-##      column names of 'coord'
-
-## It is always a good idea to center the "covariates" here
-## coord[,1:2] because the parameters are "more orthogonal"
-coord <- scale(coord, scale = FALSE)
-
-## Compare the two following lines of code and comment
-M0 <- fitspatgev(rain, coord[,1:2], loc.form, scale.form, shape.form)
-M1 <- fitspatgev(rain, scale(coord[,1:2]), loc.form, scale.form, shape.form)
-
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Show/code.R

@@ -1,44 +1,0 @@
-rm(list=ls())
-library(SpatialExtremes)
-
-## Illustration of a max-stable process
-set.seed(1234)
-n.sim <- 50
-coord <- seq(-5, 5, length = 1000)
-Y <- sqrt(2 * pi) * pmax(rgp(n.sim, coord, "powexp", range = 1, smooth = 1.9999), 0)
-poisson <- 1 / cumsum(rexp(n.sim))
-
-Phi <- poisson * Y
-Z <- apply(Phi, 2, max)
-
-par(mar = c(4, 5, 0.5, 0.5), ps = 14)
-matplot(coord, t(Phi), lty = 1, col = "grey", ylab = expression(Z(s)), xlab = expression(s),
-        type = "l")
-lines(coord, Z, col = "orange", lwd = 1.5)
-
-## Illustration of its inner structure
-observed.at <- c(-4, -1, 4)
-idx <- NULL
-for (pos in observed.at)
-    idx <- c(idx, which.min(abs(coord - pos))[1])
-
-idx.extremal.fct <- NULL
-for (j in idx)
-    idx.extremal.fct <- c(idx.extremal.fct, which(Phi[,j] == Z[j]))
-
-plot(coord, Z, col = "orange", lwd = 5,
-     xlab = expression(s), ylab = expression(Z(s)), type = "l",
-     ylim = c(0, max(Phi)))
-
-points(observed.at, Z[idx], pch = 15, col = "seagreen3", cex = 2)
-
-matlines(coord, t(Phi[idx.extremal.fct,]), lty = 1, lwd = 1.5,
-         col = c("red", "black", "blue"))
-
-matlines(coord, t(Phi[-idx.extremal.fct,]), lty = 1, col = "grey")
-
-
-
-
-
-

### 2022-03-16 Practicals/practicalSpatialExtremes/Show/_region_.tex
unzip spatial practicals

--- 2022-03-16 Practicals/practicalSpatialExtremes/Show/showv2.tex
@@
-\documentclass[mode=present,style=simple,paper=screen]{powerdot}
+\message{ !name(showv2.tex)}\documentclass[mode=present,style=simple]{powerdot}
 
+%\usepackage{simpsons}
 \usepackage[utf8]{inputenc}
-\usepackage[square]{natbib}
-\usepackage{bibentry}
+\usepackage{natbib}
 \usepackage{graphicx}
@@
 \newcommand{\advice}[1]{\HandPencilLeft{} \emph{#1}}
-\newcommand{\reminder}[1]{%
-  \fcolorbox{blue}{white}{
-    \begin{minipage}[l]{0.2\linewidth}
-      \includegraphics[width=1\textwidth]{Figures/reminder}
-    \end{minipage}
-    \begin{minipage}[r]{0.8\linewidth}
-      #1
-    \end{minipage}
-  }
-}
 
@@
   logopos={.99\slidewidth,.99\slideheight},
-  logocmd={},
+  logocmd={},%\includegraphics[height=.08\slideheight]{Figures/logoI3M}},
   counters={thm,lem,prop,cor,defn,conj,exmp,case,figure},
@@
 
-\maketitle
-
-\begin{wideslide}[toc=]{Rationale for the \texttt{SpatialExtremes} package}
-  \begin{quotation}
-    ``The aim of the \texttt{SpatialExtremes} package is to provide
-    tools for the areal modelling of extreme events. The modelling
-    strategies heavily rely on the \textcolor{blue}{extreme value
-      theory} and in particular \textcolor{blue}{block maxima}
-    techniques---unless explicitly stated.''
-  \end{quotation}
-
-  As a consequence, most often
-  \begin{itemize}
-  \item the data used by the package \textcolor{blue}{have to be
-      extreme}---do not pass daily values for instance;
-  \item \textcolor{blue}{the marginal distribution family is fixed},
-    i.e., the generalized extreme value distribution family, but you
-    have hands on how within this family parameters change in space;
-  \item \textcolor{blue}{the process family is fixed}, i.e.,
-    max-stable processes, but you have hands on which type of
-    max-stable processes to use.
-  \end{itemize}
-\end{wideslide}
-
-\section[toc=0. Max-stable process]{0. About the inner structure of max-stable processes}
-
-\begin{slide}{Spectral characterization}
-  \vspace*{-1.75em}
-  \onslide*{1}{
-    \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
-    \end{equation*}
-    where $\{\zeta_i\colon i \geq 1\}$ is a Poisson point process on
-    $(0, \infty)$ with intensity measure $\mbox{d$\Lambda$}(\zeta) =
-    \zeta^{-2} \mbox{d$\zeta$}$ and $Y_i$ independent copies of a
-    (non-negative) stochastic process such that $\mathbb{E}\{Y(s)_+\}= 1$
-    for all $s \in \mathcal{X}$.
-  }%
-  \onslide*{2-}{
-    \begin{equation*}
-      Z(s) = \max_{\varphi \in \Phi} \varphi(s), \qquad s \in \mathcal{X},
-    \end{equation*}
-    where $\Phi = \{\varphi_i\colon i \geq 1\}$ is a Poisson point
-    process on $\mathbb{C}_0$ with an appropriate intensity measure.\vspace*{-0.5em}
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.95\textwidth]{Figures/IllustMaxStable}
-    \end{figure}
-  }
-\end{slide}
-
-\begin{wideslide}{Extremal functions}
-  \vspace*{-2em}
-  \begin{figure}
-      \centering
-      \onslide*{1}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-1}}%
-      \onslide*{2}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-2}}%
-      \onslide*{3}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-3}}%
-      \onslide*{4-}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-4}}%
-    \end{figure}
-    \vspace*{-2.75em}
-    \begin{itemize}
-    \item<3-> Hidden are the random functions
-      $\Phi^+ = \{\varphi_1^+, \varphi_2^+, \ldots, \varphi_k^+\}$ of
-      $\Phi$ such that
-      \begin{equation*}
-        \varphi_j^+(s_j) = Z(s_j), \qquad j = 1, \ldots, k, \qquad
-        \text{\textcolor{red}{(extremal functions)}},
-      \end{equation*}\vspace*{-1.5em}
-    \item<4-> and the random functions $\varphi^- \in \Phi \setminus \Phi^+$,
-      i.e., satisfying
-      \begin{equation*}
-        \varphi^-(s_j) < Z(s_j), \qquad j = 1, \ldots, k, \qquad
-        \text{\textcolor{red}{(sub-extremal functions)}}
-      \end{equation*}
-    \end{itemize}
-\end{wideslide}
-
+\message{ !name(showv2.tex) !offset(76) }
 \section{1. Data and descriptive analysis}
@@
   Before introducing more advanced stuff, let's talk about data
-  format. It is pretty simple: 
+  format. It is pretty simple
   \begin{description}
@@
   might want to use additional covariates. The latter can be of two
-  types:
+  types
   \begin{description}
@@
   \begin{itemize}
-  \item As usual, you should first \textcolor{blue}{scrutinize your
+  \item As usual, you first have to \textcolor{blue}{scrutinize your
       data} (weird values, encoding of missing values, check out
-    factors, \ldots). But you always do that, don't you?
+    factors, \ldots). But you're used to that, aren't you?
   \item We focus on extremes, so you may wonder
@@
     \item are my data \textcolor{blue}{extremes}, i.e., block maxima?
-    \item is my \textcolor{blue}{block size} appropriate?
+    \item is my \textcolor{blue}{block size} relevant?
     \item what about \textcolor{blue}{seasonality}? Refine the block
-      or use a temporal covariate?
+      or use temporal covariate?
     \end{itemize}
@@
     \texttt{evd} package + a few lines of code will do the job for you
-    (homework) \ldots 
+    (homework)
   \item This will generally be OK, but now you have to go a bit
@@
     \begin{itemize}
-    \item the spatial dependence;
-    \item and the presence/absence of any spatial trends.
+    \item the spatial dependence ;
+    \item and the presence / absence of any spatial trends.
     \end{itemize}
@@
 
-\begin{slide}[toc=Spatial dependence]{Spatial dependence \citep{Cooley2006}}
+\begin{slide}{Spatial dependence}
   \begin{itemize}
@@
 
-  \reminder{\begin{equation*}
+  \fcolorbox{blue}{white}{
+
+    \Homer
+    \begin{equation*}
       \theta(h) = - z \log \Pr\{Z(s) \leq z, Z(s + h) \leq z\}
     \end{equation*}
-    where $1 \leq \theta(h) \leq 2$, with complete dependence iff
-    $\theta(h) = 1$ and independence iff $\theta(h) = 2$.}
-\end{slide}
-
-\begin{slide}[toc=]{Spatial dependence (2) \citep{Dombry2017}}
-  \begin{itemize}
-  \item Another recent summary measure of the spatial dependence is
-    the \textcolor{blue}{extremal concurrence probability function}
-    $p\colon h \mapsto p(h) \in [0, 1]$ where
-    \begin{equation*}
-      p(h) \onslide*{3-}{\textcolor{red}{:= p(s, s+h)}} = \Pr\{\exists! \varphi \in \Phi\colon \varphi(s) = Z(s), \
-      \varphi(s+h) = Z(s+h)\},
-    \end{equation*}
-    \pause i.e., there is a single extremal function at position $s$ and
-    $s+h$.
-  \end{itemize}
-  \vfill
-  \reminder{%
-    \begin{align*}
-      p(h) &= \mathbb{E}[\mbox{sign}\{Z(s) - \tilde Z(s)\}
-             \mbox{sign}\{Z(s+h) - \tilde Z(s+h)\}]\\
-      &= \text{\textcolor{blue}{Kendall's $\tau$}},
-    \end{align*}
-    and $0 \leq p(h) \leq 1$ with complete dependence iff $p(h) =
-    1$ and independence iff $p(h) = 0$.
+    and that $1 \leq \theta(h) \leq 2$ with complete dependence iff
+    $\theta(h) = 1$ and independence iff $\theta(h) = 2$.
   }
-  \onslide{3-}{\textcolor{red}{\HandRight{} NB: here we assume a stationary dependence structure!}}
 \end{slide}
@@
 
-\begin{slide}[toc=]{The \texttt{concprob} function}
+\begin{wideslide}[toc=]{Extremal concurrence probabilities: the \texttt{concprob} function}
   \begin{itemize}
-  \item Run the file \texttt{concprob.R}. You should get the figure
-    below.
-  \item<2-> You can also use a binned version with \texttt{n.bins =
+  \item Run the file \texttt{fmadogram.R}. You should get the figure
+    below. \textcolor{blue}{Any questions?}
+  \item<2-> No? What's wrong?
+  \item<3-> You can also use a binned version with \texttt{n.bins =
       300}\ldots
@@
     \centering
-    \onslide*{1}{\includegraphics[width=.66\textwidth]{Figures/concprob}}
-    \onslide*{2-}{\includegraphics[width=.66\textwidth]{Figures/concprobBinned}}
-    \caption{Use of the \texttt{concprob} function to assess the spatial dependence.}
+    \onslide*{1-2}{\includegraphics[width=.75\textwidth]{Figures/fmadogram}}
+    \onslide*{3-}{\includegraphics[width=.75\textwidth]{Figures/fmadogramBinned}}
+    \caption{Use of the \texttt{fmadogram} function to assess the spatial dependence.}
+    \label{fig:fmadogram}
   \end{figure}
   %%%\onslide{4}{\advice{Always think about what should be your spatial coordinates}}
-\end{slide}
-
-\begin{wideslide}[toc=]{As an aside (\texttt{AsAnAside.R}) \citep{Dombry2017}}
-  \vspace*{-1.75em}
-  \begin{itemize}
-  \item We can estimate the spatial distribution of the extremal
-    concurrence probability \textcolor{blue}{w.r.t. a given weather
-      station}, e.g., plotting
-    \begin{equation*}
-      \left\{(s, p(\text{Zurich}, s))\colon s \in \mathcal{X}\right\},
-    \end{equation*}\vspace*{-1.5em}
-  \item<2-> or estimate the expected area of \textcolor{blue}{concurrence cells}
-    \begin{equation*}
-      A(s_0) =
-      \mathbb{E}\left\{\int_\mathcal{X} 1_{\{\text{$s_0$ and $s$ are
-            concurrent}\}} \mbox{d$s$} \right\} = \int_\mathcal{X}
-        p(s_0, s) \mbox{d$s$}
-    \end{equation*}
-  \end{itemize}
-  \vspace*{-1.5em}
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{Figures/concMap1}\hfill%
-    \onslide{2}{\includegraphics[width=0.45\textwidth]{Figures/concMap2}}%
-  \end{figure}
 \end{wideslide}
@@
   \begin{itemize}
-  \item We can do a \textcolor{blue}{symbol plot} see the file \texttt{SpatialTrends.R}.
+  \item We can do a \textcolor{blue}{symbol plot} but the package
+    doesn't have (yet?) a function for this---mainly because it's
+    application specific.
+  \item Examples at \texttt{SpatialTrends.R} and
+    \texttt{SpatialTrends2.R}
   \end{itemize}
@@
     \onslide*{3-}{\includegraphics[width=\textwidth]{Figures/symbolPlot2}}
-    \caption{Symbol plot for the Swiss precipitation data.}
-  \end{figure}\vfill
+    \caption{Symbol plot for the swiss precipitation data.}
+  \end{figure}
   \onslide*{4-}{\advice{When exporting figures into eps/pdf, always
@@
   \item The data exhibit some \textcolor{blue}{spatial
-      dependence} and there is still some (weak) dependance at a
-    separation lag of 100km.
+      dependence}. The extremal coefficient is around 1.7 for a
+    separation lag of 100km---extremes are still not independent but
+    close to.
   \item There's a clear \textcolor{blue}{north-west / south-east
       gradient} in the intensities of rainfall storms.
-  \item In conclusion it makes sense to use max-stable
-    \textcolor{blue}{processes} whose \textcolor{blue}{marginal
-      parameters are not constant across space}.
-  \item More specifically, we have:
+  \item In conclusion it makes sense to use max-stable models whose
+    \textcolor{blue}{marginal parameters are not constant across
+      space}.
+  \item More specifically, we have
     \begin{itemize}
@@
 
-\section{2. Simple max-stable processes}
-
-\begin{slide}{Max-stable models}
+\begin{slide}{Homework}
   \begin{itemize}
-  \item In this section we \textcolor{blue}{focus only on the spatial
-      dependence} and so assume that the margins are known and
-    \textcolor{blue}{unit Fréchet}---this is a standard choice in
-    extreme value theory.
-  \item From the spectral characterization
-    \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
-    \end{equation*}
-    we can propose several parametric models for spatial
-    extremes. Hence by letting $Y$ to be
-    \begin{description}
-    \item[Gaussian densities] with random displacements we get the
-      \textcolor{blue}{Smith} process;
-    \item[Gaussian] we get the \textcolor{blue}{Schlather} process;
-    \item[Log-normal] (with a drift) we get the \textcolor{blue}{Brown--Resnick} process;
-    \item[Gaussian] but elevated to some power we get the
-      \textcolor{blue}{Extremal-$t$} process.
-    \end{description}
+  \item Have a look at the temperature and the wind gust data;
+  \item Do a descriptive analysis for these two data sets.
   \end{itemize}
@@
 
-\begin{slide}[toc=]{Dependence parameters}
-  \begin{description}
-  \item[Smith] Elements of the covariance matrix appearing in the
-    Gaussian densities;
-  \item[Schlather] Parameters of the correlation function;
-  \item[Brown--Resnick] Parameters of the semi-variogram;
-  \item[Extremal--$t$] Parameters of the correlation function and
-    degrees of freedom.
-  \end{description}
-  \begin{itemize}
-  \item Since the margins are fixed, we only need to get estimates for
-    the dependence parameters.
-  \item How can we do that?
-  \end{itemize}
-\end{slide}
-
-\begin{wideslide}[toc=Least squares,method=file]{Least squares
-    (\texttt{leastSquares.R}) \citep{Smith1990}}
-  \vspace*{-0.5em}
-  \begin{equation*}
-    \underset{\psi \in \Psi}{\arg \min} \sum_{1 \leq i < j \leq k}
-    \left\{\theta(s_j - s_j; \psi) - \hat \theta(s_i - s_j)
-    \right\}^2,
-  \end{equation*}
-  where $\theta(\cdot; \psi)$ is the extremal coefficient obtained
-  from the max-stable model with dependence parameters set to $\psi$
-  and $\hat \theta(\cdot)$ is any empirical estimates of the extremal
-  coefficient, e.g., $F$--madogram based.
-
-  \begin{minipage}[l]{.33\linewidth}
-{\tiny
-\begin{verbatim}
-> M0
-        Estimator: Least Squares
-            Model: Schlather
-         Weighted: TRUE
-  Objective Value: 3592.429
-Covariance Family: Whittle-Matern
-
-Estimates
-  Marginal Parameters:
-  Assuming unit Frechet.
-
-  Dependence Parameters:
-  range   smooth
-54.3239   0.4026
-
-Optimization Information
-  Convergence: successful
-  Function Evaluations: 61
-\end{verbatim}
-}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[r]{.6\linewidth}
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.49\textwidth]{Figures/empExtCoeffRain}%
-      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/leastSquares}}
-      \caption{Fitting simple max-stable processes from least
-        squares.}
-      \label{fig:leastSquares}
-    \end{figure}
-  \end{minipage}
-\end{wideslide}
-
-\begin{wideslide}[toc=Pairwise likelihood,method=file]{Pairwise
-    likelihood (\texttt{pairwiseLlik.R}) \citep{Padoan2010}}
-  \vspace*{-2em}
-  \begin{equation*}
-    \underset{\psi \in \Psi}{\arg \max} \sum_{\ell = 1}^n \sum_{1 \leq
-      i < j \leq k} \log f\{z_\ell(s_i), z_\ell(s_j); \psi\},
-  \end{equation*}
-  where $f(\cdot, \cdot; \psi)$ is the bivariate density of the
-  considered max-stable model.
-
-    \begin{minipage}[l]{.33\linewidth}
-      {\tiny
-\begin{verbatim}
-        Estimator: MPLE
-            Model: Schlather
-         Weighted: FALSE
-   Pair. Deviance: 1136863
-              TIC: 1137456
-Covariance Family: Whittle-Matern
-
-Estimates
-  Marginal Parameters:
-  Assuming unit Frechet.
-
-  Dependence Parameters:
-  range   smooth
-50.1976   0.3713
-
-Standard Errors
-  range   smooth
-20.7085   0.0789
-
-Asymptotic Variance Covariance
-        range       smooth
-range   428.841018   -1.570081
-smooth   -1.570081    0.006225
-Optimization Information
-  Convergence: successful
-  Function Evaluations: 67
-\end{verbatim}
-}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[r]{.6\linewidth}
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.49\textwidth]{Figures/empExtCoeffRain}%
-      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/pairwiseLlik}}
-      \caption{Fitting simple max-stable processes maximizing pairwise
-        likelihood.}
-      \label{fig:pairLlik}
-    \end{figure}
-  \end{minipage}
-\end{wideslide}
-
-\begin{slide}[method=direct,toc= Model selection]{Model Selection \citep{Varin2005}}
-  \begin{itemize}
-  \item The advantage of the pairwise likelihood estimator over the
-    least squares one is that you can do \textcolor{blue}{model
-      selection}.
-  \item For instance one can use the ${\rm TIC}$,
-    \textcolor{blue}{Takeuchi Information Criterion} or sometimes
-    known as ${\rm CLIC}$, Composite Likelihood Information Criterion,
-    \begin{equation*}
-      {\rm TIC} = 2 \ell_{\rm pairwise}(\hat \psi) - 2
-      \mbox{tr}\{J(\hat \psi) H^{-1} (\hat \psi)\},
-    \end{equation*}
-    $H(\hat\psi) = \mathbb{E} \{\nabla^2 \ell_{\rm pairwise}(Y; \hat
-    \psi) \}$, $J(\hat \psi) = \mbox{Var}\{\nabla \ell_{\rm
-      pairwise}(Y; \hat\psi)\}$.
-  \item From our previous fitted models, we get
-{\tiny
-\begin{verbatim}
-> TIC(M0,M1,M2)
-     M1      M2      M0
-1133660 1134823 1137449
-\end{verbatim}
-}
-  \end{itemize}
-\end{slide}
-
-\begin{slide}[toc=Simulation,method=direct]{Simulating simple max-stable processes
-    (\texttt{simulation.R}) \citep{Schlather2002,Dombry2016}}
-  \begin{itemize}
-  \item Once you have fitted a suitable model, you usually want to
-    simulate from it.
-  \item Simulation from max-stable models is rather complex, recall
-    that
-    \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X}.
-    \end{equation*}
-  \end{itemize}
-  \begin{minipage}[l]{.6\linewidth}
-{\tiny
-\begin{verbatim}
-sim <- rmaxstab(n.obs, cbind(x, y), "twhitmat", DoF = 4,
-+               nugget = 0, range = 3, smooth = 1)
-> sim
-            [,1]       [,2]      [,3]      [,4]       [,5]
- [1,]  3.8048914  0.4767980 6.3613989 1.4548317  1.0433912
- [2,]  1.2200332  0.6711422 0.8078701 2.0928629  0.7537061
- [3,]  0.5466466  2.0498561 4.8852572 2.3497976  0.6857268
- [4,]  ...
-\end{verbatim}
-}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[r]{.33\linewidth}
-    \begin{figure}
-      \centering
-      \includegraphics[width=\textwidth]{Figures/simulation}
-      \caption{One simulation on a 50 x 50 grid from the extremal--$t$
-        model. (log scale)}
-    \end{figure}
-  \end{minipage}
-\end{slide}
-
-\begin{slide}[toc=Debrief \#2]{What we have learned so far (apart from
-    using \texttt{SpatialExtremes})}
-  \begin{itemize}
-  \item The Smith model is clearly not a sensible model for our
-    data---because of its linear behaviour near the origin;
-  \item Schlather, Brown--Resnick and Extremal--$t$ seems relevant;
-  \item According to the ${\rm TIC}$, the Extremal--$t$ should be
-    preferred.
-  \end{itemize}
-\end{slide}
-
-\section{3. Trends surfaces}
-
-\begin{slide}[toc=]{From generalized extreme value margins to unit
-    Fréchet ones}
-  \begin{itemize}
-  \item Alright! We are able to handle the spatial dependence, but
-    \textcolor{blue}{we assume that our data have unit Fréchet
-      margins}. This is not realistic at all!
-  \item Fortunately, if $Y \sim {\rm GEV}(\mu, \sigma, \xi)$ then
-    \begin{equation*}
-      Z = \left(1 + \xi \frac{Y - \mu}{\sigma} \right)^{1 / \xi} \sim
-      \text{Unit Fréchet}=\mbox{GEV}(1,1,1).
-    \end{equation*}
-    \pause
-  \item And since we are extreme value and spatial guys
-    \begin{equation*}
-      Z\textcolor{blue}{(s)} = \left\{1 + \xi\textcolor{blue}{(s)}
-        \frac{Y\textcolor{blue}{(s)} -
-          \mu\textcolor{blue}{(s)}}{\sigma\textcolor{blue}{(s)}}
-      \right\}^{1 / \xi\textcolor{blue}{(s)}}, \qquad s \in \mathcal{X},
-    \end{equation*}
-    is a simple max-stable process.
-  \item Hence we can use the maximum pairwise likelihood estimator as
-    before---up to an \textcolor{blue}{additional Jacobian term}.
-  \end{itemize}
-\end{slide}
-
-\begin{slide}[toc=Spatial GEV]{Omitting the spatial dependence}
-  \begin{itemize}
-  \item With simple max-stable models, we omitted the marginal parameters.
-  \item Here we will \textcolor{blue}{omit the spatial dependence for
-      a while} and consider locations as being
-    \textcolor{blue}{mutually independent}, i.e., use independence
-    likelihood
-    \begin{equation*}
-      \underset{\psi \in \Psi}{\arg \max} \sum_{i=1}^k \ell_{\rm
-        GEV}\{y(s_i); \psi\}.
-    \end{equation*}
-  \item This is a kind of ``\emph{spatial GEV}'' where $\psi$ is a
-    vector of marginal parameters.
-  \end{itemize}
-\end{slide}
-
-\begin{slide}[toc=,method=direct]{Defining trend surfaces}
-  \vspace*{-1em}
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/symbolPlot2}
-    \caption{Symbol plot for the Swiss precipitation data.}
-  \end{figure}
-  This suggests that
-  \begin{align*}
-    \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
-      \mu} {\rm lat}(x) + \beta_{3, \mu} {\rm lon}(s) \times {\rm lat}(s),\\
-    \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
-      \sigma} {\rm lat}(s) + \beta_{3, \sigma} {\rm lon}(s) \times {\rm lat}(s),\\
-    \xi(s) &= \beta_{0,\xi},
-  \end{align*}
-  or equivalently with the \texttt{R} language\vspace*{-0.5em}
-{\small
-\begin{verbatim}
-loc.form <- scale.form <- y ~ lon * lat; shape.form <- y ~ 1
-\end{verbatim}
-}
-\end{slide}
-
-\begin{slide}[toc=,method=direct]{Fitting the \emph{spatial GEV} model
-    (\texttt{spatialGEV.R}) \citep{Davison2012}}
-  \vspace*{-1em}
-  \tiny{
-\begin{verbatim}
-      Model: Spatial GEV model
-   Deviance: 29303.81
-        TIC: 29499.38
-
-    Location Parameters:
-locCoeff1  locCoeff2  locCoeff3  locCoeff4
-   27.132      1.846     -3.656     -1.080
-       Scale Parameters:
-scaleCoeff1  scaleCoeff2  scaleCoeff3  scaleCoeff4
-     9.7850       0.7023      -1.0858      -0.5531
-       Shape Parameters:
-shapeCoeff1
-     0.1572
-
-Standard Errors
-  locCoeff1    locCoeff2    locCoeff3    locCoeff4  scaleCoeff1  scaleCoeff2
-    1.13326      0.34864      0.45216      0.38361      0.76484      0.28446
-scaleCoeff3  scaleCoeff4  shapeCoeff1
-    0.31267      0.27566      0.05878
-
-Asymptotic Variance Covariance
-             locCoeff1   locCoeff2   locCoeff3   locCoeff4   scaleCoeff1
-locCoeff1     1.2842711   0.1131400  -0.1740921  -0.0729564   0.6570988
-locCoeff2     0.1131400   0.1215498  -0.0623759   0.0149596   0.0521630
-locCoeff3    -0.1740921  -0.0623759   0.2044448   0.0576622  -0.1086629
-locCoeff4    -0.0729564   0.0149596   0.0576622   0.1471593  -0.0346376
-scaleCoeff1   0.6570988   0.0521630  -0.1086629  -0.0346376   0.5849729
-...
-
-Optimization Information
-  Convergence: successful
-  Function Evaluations: 2135
-\end{verbatim}
-    }
-\end{slide}
-
-\begin{wideslide}[toc=Prediction \#1]{Get predictions (\texttt{predictionSpatialGEV.R})}
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{Figures/symbolPlot}%
-    \includegraphics[width=.5\textwidth]{Figures/predictionSpatialGEV}
-    \caption{Left: symbol plot. Right: Prediction of the pointwise
-      25-year return levels from a fitted spatial GEV model.}
-    \label{fig:prediction}
-  \end{figure}
-  \pause
-  \begin{itemize}
-  \item But don't we forget something??? \pause
-  \item Model selection?
-  \end{itemize}
-\end{wideslide}
-
-\begin{slide}[toc=Model selection \#2,method=direct]{Model selection \#2
-    (\texttt{modelSelection.R}) \citep{Chandler2007,Kent1982}}
-  \begin{itemize}
-  \item Typically here we would like to test if a given covariate is
-    required or not
-  \item Hence we're dealing with nested model for which
-    \textcolor{blue}{composite likelihood ratio test} are especially
-    suited
-    \begin{equation*}
-      2 \{\ell_{\rm composite}(\hat \psi) - \ell_{\rm composite}(\hat
-      \phi_{\lambda_0}, \lambda_0)\} \longrightarrow \sum_{j=1}^p
-      \lambda_j X_i, \qquad n \to \infty.
-    \end{equation*}
-  \end{itemize}
-{\tiny
-\begin{verbatim}
-Eigenvalue(s): 2.7 1.95
-
-Analysis of Variance Table
-   MDf Deviance Df  Chisq Pr(> sum lambda Chisq)
-M2   7    29328
-M0   9    29306  2 22.265               0.008273 **
----
-Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
-\end{verbatim}
-}
-\advice{Always check that your models are nested. The code won't do
-  that for you!}
-\end{slide}
-
-\begin{slide}[toc=Debrief \#3]{What we have learned so far (apart from
-    using \texttt{SpatialExtremes})}
-  \begin{itemize}
-  \item Based on the \texttt{spatial GEV} model, we identify what
-    seems to be relevant trend surfaces for the marginal parameters:
-    \begin{align*}
-      \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
-        \mu} {\rm lat}(s) + \beta_{3,\mu} {\rm lon}(s) {\rm lat(s)},\\
-      \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
-        \sigma} {\rm lat}(s),\\
-      \xi(s) &= \beta_{0,\xi},
-    \end{align*}
-  \end{itemize}
-\end{slide}
-
-\section{4. General max-stable processes}
-
-\begin{wideslide}[toc=Fitting,method=direct]{Fitting a max-stable process with trend
-    surfaces}
-  \begin{itemize}
-  \item Now it's time to combine everything, i.e.,
-    \textcolor{blue}{trend surfaces + dependence}.
-  \item The syntax won't be a big surprise
-  \end{itemize}
-  {\tiny
-\begin{verbatim}
-M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form, scale.form, shape.form)
-
-       Estimator: MPLE
-            Model: Extremal-t
-         Weighted: FALSE
-   Pair. Deviance: 2237562
-              TIC: 2249206
-Covariance Family: Whittle-Matern
-
-Estimates
-  Marginal Parameters:
-    Location Parameters:
-locCoeff1  locCoeff2  locCoeff3  locCoeff4
-27.136295   0.060145  -0.164755  -0.001117
-       Scale Parameters:
-scaleCoeff1  scaleCoeff2  scaleCoeff3
-    9.88857      0.02869     -0.04581
-       Shape Parameters:
-shapeCoeff1
-     0.1727
-  Dependence Parameters:
-   range    smooth       DoF
-225.9452    0.3645    4.1566
-...
-\end{verbatim}
-}
-\end{wideslide}
-
-\begin{wideslide}[method=direct]{Model checking \citep{Davison2012}}
-  \begin{minipage}[l]{.49\linewidth}
-    \begin{itemize}
-    \item When you want to check your fitted max-stable model, you
-      usually want to check if
-      \begin{itemize}
-      \item observations at each single location are well modelled:
-        \textcolor{blue}{return level plot};
-      \item the dependence is well captured: \textcolor{blue}{extremal
-          coefficient function}.
-      \end{itemize}
-    \item This can be done using a single line of code
-\begin{verbatim}
-> plot(M0)
-\end{verbatim}
-    \end{itemize}
-  \end{minipage}
-  \hfill
-  \onslide{2}{
-  \begin{minipage}[r]{.49\linewidth}
-    \begin{figure}
-      \centering
-      \includegraphics[width=\textwidth]{Figures/checkFinalModel}
-      \caption{Model checking for a fitted max-stable process having
-        trend surfaces.}
-    \end{figure}
-  \end{minipage}
-}
-\end{wideslide}
-
-\begin{wideslide}{Predictions (\texttt{simulationFinal.R})}
-  \begin{itemize}
-  \item Prediction works as for the \emph{spatial GEV model} thanks to
-    the \texttt{predict} function.
-  \item But beware these predictions are pointwise---no spatial
-    dependence at all!!!
-  \item If you want to do take into account spatial dependence then
-    you need to simulate from your fitted model.
-  \end{itemize}
-  \vspace*{-2em}
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{Figures/oneSimFinalModel}
-    \caption{One simulation from our fitted extremal--$t$ model with trend surfaces.}
-  \end{figure}
-\end{wideslide}
-
-\section{5. Conclusion}
-
-\begin{wideslide}[toc=]{What we haven't seen}
-  \begin{itemize}
-  \item Using weighted pairwise likelihood;
-  \item Many (many!) utility functions. Highly recommended to have a
-    look at the documentation;
-  \item The package has a vignette: \texttt{vignette("SpatialExtremesGuide")};
-  \item Copula models---although I do not recommend their use for
-    spatial extremes;
-  \item Bayesian hierarchical models;
-  \item Unconditional simulations: several implementations (including
-    exact simulations)
-  \item Conditional simulations---really CPU demanding.
-  \end{itemize}
-  \pause
-  \HandRight{} A rather recent review on max-stable processes with R
-  code is given by \citet{Ribatet2013b}
-  \vfill
-  \begin{center}
-    \Large THANK YOU!
-  \end{center}
-  \vfill
-\end{wideslide}
-
-\begin{wideslide}{References}
-  \vspace*{-1.5em}
-  {\footnotesize
-  \bibliographystyle{apalike}
-  \nobibliography{/Users/Mathieu/Dropbox/biblio_ribatet}
-
-  \begin{itemize}
-  \item[\checkmark] \bibentry{Chandler2007}
-  \item[\checkmark] \bibentry{Cooley2006}
-  \item[\checkmark] \bibentry{Davison2012}
-  \item[\checkmark] \bibentry{Dombry2016}
-  \item[\checkmark] \bibentry{Dombry2017}
-  \item[\checkmark] \bibentry{Padoan2010}
-  \item[\checkmark] \bibentry{Ribatet2013b}
-  \item[\checkmark] \bibentry{Schlather2002}
-  \item[\checkmark] \bibentry{Smith1990}
-  \item[\checkmark] \bibentry{Varin2005}
-  \end{itemize}
-  }
-\end{wideslide}
-
-%\bibliography{/Users/Mathieu/Dropbox/biblio_ribatet}
-%\bibliographystyle{apalike}
+\message{ !name(showv2.tex) !offset(601) }
 

### 2022-03-16 Practicals/practicalSpatialExtremes/Show/show.tex
unzip spatial practicals

--- 2022-03-16 Practicals/practicalSpatialExtremes/Show/showv2.tex
@@
-\documentclass[mode=present,style=simple,paper=screen]{powerdot}
+\documentclass[blackslide,style=simple]{powerdot}
 
 \usepackage[utf8]{inputenc}
-\usepackage[square]{natbib}
-\usepackage{bibentry}
+\usepackage{natbib}
 \usepackage{graphicx}
@@
 \newcommand{\advice}[1]{\HandPencilLeft{} \emph{#1}}
-\newcommand{\reminder}[1]{%
-  \fcolorbox{blue}{white}{
-    \begin{minipage}[l]{0.2\linewidth}
-      \includegraphics[width=1\textwidth]{Figures/reminder}
-    \end{minipage}
-    \begin{minipage}[r]{0.8\linewidth}
-      #1
-    \end{minipage}
-  }
-}
 
-\title{Statistical modelling of spatial extremes using the SpatialExtremes package}
+\title{Modelling spatial extremes with the SpatialExtremes package}
 \author{M. Ribatet}
-\date{University of Montpellier}
+\date{Pan-American Advanced Study Institute on Spatio-Temporal Statistics}
 % \date{
@@
 \pdsetup{
-  lf={The \texttt{SpatialExtremes} package},
+  lf=The \texttt{SpatialExtremes} package,
   rf={Mathieu Ribatet},
@@
 
-\section[toc=0. Max-stable process]{0. About the inner structure of max-stable processes}
-
-\begin{slide}{Spectral characterization}
-  \vspace*{-1.75em}
-  \onslide*{1}{
-    \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
-    \end{equation*}
-    where $\{\zeta_i\colon i \geq 1\}$ is a Poisson point process on
-    $(0, \infty)$ with intensity measure $\mbox{d$\Lambda$}(\zeta) =
-    \zeta^{-2} \mbox{d$\zeta$}$ and $Y_i$ independent copies of a
-    (non-negative) stochastic process such that $\mathbb{E}\{Y(s)_+\}= 1$
-    for all $s \in \mathcal{X}$.
-  }%
-  \onslide*{2-}{
-    \begin{equation*}
-      Z(s) = \max_{\varphi \in \Phi} \varphi(s), \qquad s \in \mathcal{X},
-    \end{equation*}
-    where $\Phi = \{\varphi_i\colon i \geq 1\}$ is a Poisson point
-    process on $\mathbb{C}_0$ with an appropriate intensity measure.\vspace*{-0.5em}
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.95\textwidth]{Figures/IllustMaxStable}
-    \end{figure}
-  }
-\end{slide}
-
-\begin{wideslide}{Extremal functions}
-  \vspace*{-2em}
-  \begin{figure}
-      \centering
-      \onslide*{1}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-1}}%
-      \onslide*{2}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-2}}%
-      \onslide*{3}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-3}}%
-      \onslide*{4-}{\includegraphics[width=0.7\textwidth]{Figures/IllustExtremalFct-4}}%
-    \end{figure}
-    \vspace*{-2.75em}
-    \begin{itemize}
-    \item<3-> Hidden are the random functions
-      $\Phi^+ = \{\varphi_1^+, \varphi_2^+, \ldots, \varphi_k^+\}$ of
-      $\Phi$ such that
-      \begin{equation*}
-        \varphi_j^+(s_j) = Z(s_j), \qquad j = 1, \ldots, k, \qquad
-        \text{\textcolor{red}{(extremal functions)}},
-      \end{equation*}\vspace*{-1.5em}
-    \item<4-> and the random functions $\varphi^- \in \Phi \setminus \Phi^+$,
-      i.e., satisfying
-      \begin{equation*}
-        \varphi^-(s_j) < Z(s_j), \qquad j = 1, \ldots, k, \qquad
-        \text{\textcolor{red}{(sub-extremal functions)}}
-      \end{equation*}
-    \end{itemize}
-\end{wideslide}
-
 \section{1. Data and descriptive analysis}
@@
     \end{equation*}
+  \item Recall that $1 \leq \theta(h) \leq 2$ where complete
+    dependence iff $\theta(h) = 1$ and independence iff $\theta(h) =
+    2$.
   \item The \texttt{fmadogram} function will estimate (empirically)
@@
   \end{itemize}
-
-  \reminder{\begin{equation*}
-      \theta(h) = - z \log \Pr\{Z(s) \leq z, Z(s + h) \leq z\}
-    \end{equation*}
-    where $1 \leq \theta(h) \leq 2$, with complete dependence iff
-    $\theta(h) = 1$ and independence iff $\theta(h) = 2$.}
-\end{slide}
-
-\begin{slide}[toc=]{Spatial dependence (2) \citep{Dombry2017}}
-  \begin{itemize}
-  \item Another recent summary measure of the spatial dependence is
-    the \textcolor{blue}{extremal concurrence probability function}
-    $p\colon h \mapsto p(h) \in [0, 1]$ where
-    \begin{equation*}
-      p(h) \onslide*{3-}{\textcolor{red}{:= p(s, s+h)}} = \Pr\{\exists! \varphi \in \Phi\colon \varphi(s) = Z(s), \
-      \varphi(s+h) = Z(s+h)\},
-    \end{equation*}
-    \pause i.e., there is a single extremal function at position $s$ and
-    $s+h$.
-  \end{itemize}
-  \vfill
-  \reminder{%
-    \begin{align*}
-      p(h) &= \mathbb{E}[\mbox{sign}\{Z(s) - \tilde Z(s)\}
-             \mbox{sign}\{Z(s+h) - \tilde Z(s+h)\}]\\
-      &= \text{\textcolor{blue}{Kendall's $\tau$}},
-    \end{align*}
-    and $0 \leq p(h) \leq 1$ with complete dependence iff $p(h) =
-    1$ and independence iff $p(h) = 0$.
-  }
-  \onslide{3-}{\textcolor{red}{\HandRight{} NB: here we assume a stationary dependence structure!}}
 \end{slide}
@@
   \item Run the file \texttt{fmadogram.R}. You should get the figure
-    below.
-  \item<2-> You can also use a binned version with \texttt{n.bins =
+    below. \textcolor{blue}{Any questions?}
+  \item<2-> No? What's wrong?
+  \item<3-> You can also use a binned version with \texttt{n.bins =
       300}\ldots
@@
     \centering
-    \onslide*{1}{\includegraphics[width=.75\textwidth]{Figures/fmadogram}}
-    \onslide*{2-}{\includegraphics[width=.75\textwidth]{Figures/fmadogramBinned}}
+    \onslide*{1-2}{\includegraphics[width=.75\textwidth]{Figures/fmadogram}}
+    \onslide*{3-}{\includegraphics[width=.75\textwidth]{Figures/fmadogramBinned}}
     \caption{Use of the \texttt{fmadogram} function to assess the spatial dependence.}
@@
 
-\begin{slide}[toc=]{The \texttt{concprob} function}
-  \begin{itemize}
-  \item Run the file \texttt{concprob.R}. You should get the figure
-    below.
-  \item<2-> You can also use a binned version with \texttt{n.bins =
-      300}\ldots
-  \end{itemize}
-  \begin{figure}
-    \centering
-    \onslide*{1}{\includegraphics[width=.66\textwidth]{Figures/concprob}}
-    \onslide*{2-}{\includegraphics[width=.66\textwidth]{Figures/concprobBinned}}
-    \caption{Use of the \texttt{concprob} function to assess the spatial dependence.}
-  \end{figure}
-  %%%\onslide{4}{\advice{Always think about what should be your spatial coordinates}}
-\end{slide}
-
-\begin{wideslide}[toc=]{As an aside (\texttt{AsAnAside.R}) \citep{Dombry2017}}
-  \vspace*{-1.75em}
-  \begin{itemize}
-  \item We can estimate the spatial distribution of the extremal
-    concurrence probability \textcolor{blue}{w.r.t. a given weather
-      station}, e.g., plotting
-    \begin{equation*}
-      \left\{(s, p(\text{Zurich}, s))\colon s \in \mathcal{X}\right\},
-    \end{equation*}\vspace*{-1.5em}
-  \item<2-> or estimate the expected area of \textcolor{blue}{concurrence cells}
-    \begin{equation*}
-      A(s_0) =
-      \mathbb{E}\left\{\int_\mathcal{X} 1_{\{\text{$s_0$ and $s$ are
-            concurrent}\}} \mbox{d$s$} \right\} = \int_\mathcal{X}
-        p(s_0, s) \mbox{d$s$}
-    \end{equation*}
-  \end{itemize}
-  \vspace*{-1.5em}
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{Figures/concMap1}\hfill%
-    \onslide{2}{\includegraphics[width=0.45\textwidth]{Figures/concMap2}}%
-  \end{figure}
-\end{wideslide}
-
 \begin{wideslide}[toc=Spatial trends]{Spatial trends}
@@
 
+\begin{slide}{Homework}
+  \begin{itemize}
+  \item Have a look at the temperature and the wind gust data;
+  \item Do a descriptive analysis for these two data sets.
+  \end{itemize}
+\end{slide}
+
 \section{2. Simple max-stable processes}
+\label{sec:simple-max-stable}
 
@@
     extreme value theory.
-  \item From the spectral characterization
+  \item From the spectral characterization (Dan's lecture)
     \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X},
+      Z(x) = \max_{i \geq 1} \zeta_i Y_i(x), \qquad x \in \mathcal{X}
+      \subset \mathbb{R}^d,
     \end{equation*}
@@
     \item[Gaussian densities] with random displacements we get the
-      \textcolor{blue}{Smith} process;
-    \item[Gaussian] we get the \textcolor{blue}{Schlather} process;
-    \item[Log-normal] (with a drift) we get the \textcolor{blue}{Brown--Resnick} process;
-    \item[Gaussian] but elevated to some power we get the
-      \textcolor{blue}{Extremal-$t$} process.
+      Smith process;
+    \item[Gaussian] we get the Schlather process;
+    \item[Log-normal] (with a drift) we get the Brown--Resnick process;
+    \item[Gaussian] but elevated to some power we get the Extremal-$t$
+      process.
     \end{description}
@@
 
-\begin{wideslide}[toc=Least squares,method=file]{Least squares
-    (\texttt{leastSquares.R}) \citep{Smith1990}}
+\begin{wideslide}[toc=Least squares,method=file]{Least squares (\texttt{leastSquares.R})}
   \vspace*{-0.5em}
@@
     \underset{\psi \in \Psi}{\arg \min} \sum_{1 \leq i < j \leq k}
-    \left\{\theta(s_j - s_j; \psi) - \hat \theta(s_i - s_j)
+    \left\{\theta(x_j - x_j; \psi) - \hat \theta(x_i - x_j)
     \right\}^2,
@@
 \begin{wideslide}[toc=Pairwise likelihood,method=file]{Pairwise
-    likelihood (\texttt{pairwiseLlik.R}) \citep{Padoan2010}}
+    likelihood (\texttt{pairwiseLlik.R})}
   \vspace*{-2em}
@@
     \underset{\psi \in \Psi}{\arg \max} \sum_{\ell = 1}^n \sum_{1 \leq
-      i < j \leq k} \log f\{z_\ell(s_i), z_\ell(s_j); \psi\},
+      i < j \leq k} \log f\{z_\ell(x_i), z_\ell(x_j); \psi\},
   \end{equation*}
@@
       \includegraphics[width=0.49\textwidth]{Figures/empExtCoeffRain}%
-      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/pairwiseLlik}}
+      \onslide{2}{\includegraphics[width=0.49\textwidth]{Figures/pairwiseLLik}}
       \caption{Fitting simple max-stable processes maximizing pairwise
@@
 
-\begin{slide}[method=direct,toc= Model selection]{Model Selection \citep{Varin2005}}
+\begin{slide}[method=direct]{Model Selection}
   \begin{itemize}
@@
 \begin{verbatim}
-> TIC(M0,M1,M2)
-     M1      M2      M0
-1133660 1134823 1137449
+> TIC(M0,M1,M2,M3)
+     M2      M3      M0      M1 
+1133668 1134829 1137456 1159784 
 \end{verbatim}
@@
 \begin{slide}[toc=Simulation,method=direct]{Simulating simple max-stable processes
-    (\texttt{simulation.R}) \citep{Schlather2002,Dombry2016}}
+    (\texttt{simulation.R})}
   \begin{itemize}
@@
     \begin{equation*}
-      Z(s) = \max_{i \geq 1} \zeta_i Y_i(s), \qquad s \in \mathcal{X}.
+      Z(x) = \max_{i \geq 1} \zeta_i Y_i(x), \qquad x \in \mathcal{X}.
     \end{equation*}
@@
 
+\begin{slide}[method=direct]{Homework}
+  \begin{itemize}
+  \item Perform simulations from various max-stable processes on
+    scattered and lattice locations;
+  \item Fit a Schlather model with a powered exponential correlation
+    function;
+  \item Why do we always set \verb|nugget = 0|?
+  \item Try to put weights within the pairwise likelihood.
+  \end{itemize}
+\end{slide}
+
 \section{3. Trends surfaces}
@@
   \begin{itemize}
-  \item Alright! We are able to handle the spatial dependence, but
+  \item Alright we are able to handle the spatial dependence, but
     \textcolor{blue}{we assume that our data have unit Fréchet
@@
     \begin{equation*}
-      Z = \left(1 + \xi \frac{Y - \mu}{\sigma} \right)^{1 / \xi} \sim
-      \text{Unit Fréchet}=\mbox{GEV}(1,1,1).
+      Z = \left(1 + \xi \frac{Y - \mu}{\sigma} \right)^{1 / \xi}
     \end{equation*}
+    is a unit Fréchet random variable.
     \pause
@@
     \begin{equation*}
-      Z\textcolor{blue}{(s)} = \left\{1 + \xi\textcolor{blue}{(s)}
-        \frac{Y\textcolor{blue}{(s)} -
-          \mu\textcolor{blue}{(s)}}{\sigma\textcolor{blue}{(s)}}
-      \right\}^{1 / \xi\textcolor{blue}{(s)}}, \qquad s \in \mathcal{X},
+      Z\textcolor{blue}{(x)} = \left\{1 + \xi\textcolor{blue}{(x)}
+        \frac{Y\textcolor{blue}{(x)} -
+          \mu\textcolor{blue}{(x)}}{\sigma\textcolor{blue}{(x)}}
+      \right\}^{1 / \xi\textcolor{blue}{(x)}}, \qquad x \in \mathcal{X},
     \end{equation*}
     is a simple max-stable process.
-  \item Hence we can use the maximum pairwise likelihood estimator as
+  \item Hence we can use the pairwise likelihood estimator as
     before---up to an \textcolor{blue}{additional Jacobian term}.
@@
       \underset{\psi \in \Psi}{\arg \max} \sum_{i=1}^k \ell_{\rm
-        GEV}\{y(s_i); \psi\}.
+        GEV}\{y(x_i); \psi\}.
     \end{equation*}
@@
     \includegraphics[width=\textwidth]{Figures/symbolPlot2}
-    \caption{Symbol plot for the Swiss precipitation data.}
+    \caption{Symbol plot for the swiss precipitation data.}
   \end{figure}
-  This suggests that
+  This suggest that
   \begin{align*}
-    \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
-      \mu} {\rm lat}(x) + \beta_{3, \mu} {\rm lon}(s) \times {\rm lat}(s),\\
-    \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
-      \sigma} {\rm lat}(s) + \beta_{3, \sigma} {\rm lon}(s) \times {\rm lat}(s),\\
-    \xi(s) &= \beta_{0,\xi},
+    \mu(x) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(x) + \beta_{2,
+      \mu} {\rm lat}(x) + \beta_{3, \mu} {\rm lon}(x) \times {\rm lat}(x),\\
+    \sigma(x) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(x) + \beta_{2,
+      \sigma} {\rm lat}(x) + \beta_{3, \sigma} {\rm lon}(x) \times {\rm lat}(x),\\
+    \xi(x) &= \beta_{0,\xi},
   \end{align*}
-  or equivalently with the \texttt{R} language\vspace*{-0.5em}
-{\small
+  or equivalently with the \texttt{R} language
+{\tiny
 \begin{verbatim}
@@
 \begin{slide}[toc=,method=direct]{Fitting the \emph{spatial GEV} model
-    (\texttt{spatialGEV.R}) \citep{Davison2012}}
+    (\texttt{spatialGEV.R})}
   \vspace*{-1em}
@@
 \begin{slide}[toc=Model selection \#2,method=direct]{Model selection \#2
-    (\texttt{modelSelection.R}) \citep{Chandler2007,Kent1982}}
+    (\texttt{modelSelection.R})}
   \begin{itemize}
@@
 \begin{verbatim}
-Eigenvalue(s): 2.7 1.95
+Eigenvalue(s):
+0.06
+0.04
 
@@
    MDf Deviance Df  Chisq Pr(> sum lambda Chisq)    
-M2   7    29328
-M0   9    29306  2 22.265               0.008273 **
+M2   7    29329                                     
+M0   9    29304  2 24.924              < 2.2e-16 ***
 ---
-Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
 \end{verbatim}  
@@
     \begin{align*}
-      \mu(s) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(s) + \beta_{2,
-        \mu} {\rm lat}(s) + \beta_{3,\mu} {\rm lon}(s) {\rm lat(s)},\\
-      \sigma(s) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(s) + \beta_{2,
-        \sigma} {\rm lat}(s),\\
-      \xi(s) &= \beta_{0,\xi},
+      \mu(x) &= \beta_{0,\mu} + \beta_{1, \mu} {\rm lon}(x) + \beta_{2,
+        \mu} {\rm lat}(x),\\
+      \sigma(x) &= \beta_{0,\sigma} + \beta_{1, \sigma} {\rm lon}(x) + \beta_{2,
+        \sigma} {\rm lat}(x),\\
+      \xi(x) &= \beta_{0,\xi},
     \end{align*}
+  \end{itemize}  
+\end{slide}
+
+\begin{slide}{Homework}
+  \begin{itemize}
+  \item Produce a figure similar to Figure~\ref{fig:prediction} with
+    our best model;
+  \item Fit a \emph{spatial GEV} model where elevation appears;
+  \item Is this model appropriate?
   \end{itemize}  
@@
          Weighted: FALSE 
-   Pair. Deviance: 2237562
-              TIC: 2249206
+   Pair. Deviance: 2239596 
+              TIC: 2251000 
 Covariance Family: Whittle-Matern
@@
     Location Parameters:
-locCoeff1  locCoeff2  locCoeff3  locCoeff4
-27.136295   0.060145  -0.164755  -0.001117
+locCoeff1  locCoeff2  locCoeff3  
+ 20.65202    0.06473   -0.15630  
        Scale Parameters:
 scaleCoeff1  scaleCoeff2  scaleCoeff3  
-    9.88857      0.02869     -0.04581
+    5.25314      0.02087     -0.04029  
        Shape Parameters:
 shapeCoeff1  
-     0.1727
+     0.1892  
   Dependence Parameters:
    range    smooth       DoF  
-225.9452    0.3645    4.1566
+226.7903    0.3562    3.9615  
 ...
@@
 
-\begin{wideslide}[method=direct]{Model checking \citep{Davison2012}}
+\begin{wideslide}{Model checking}
   \begin{minipage}[l]{.49\linewidth}
@@
     \item This can be done using a single line of code
-\begin{verbatim}
-> plot(M0)
-\end{verbatim}
+      \texttt{plot(M0)}.
     \end{itemize}
@@
   \hfill
-  \onslide{2}{
+  \onslide*{2}{
   \begin{minipage}[r]{.49\linewidth}
@@
 
-\begin{wideslide}{Predictions (\texttt{simulationFinal.R})}
+\begin{wideslide}{Predictions}
   \begin{itemize}
@@
   \item If you want to do take into account spatial dependence then
-    you need to simulate from your fitted model.
+    you need to simulate from your fitted model---see
+    \texttt{simulationFinal.R}.
   \end{itemize}
@@
 
+\begin{slide}{Homework}
+  \begin{itemize}
+  \item Perform a simulation study to estimate the distribution of
+    \begin{equation*}
+      \sup_{x \in b({\rm Zurich}, 10{\rm km})} Z(x),
+    \end{equation*}
+    where $b({\rm Zurich}, 10{\rm km})$ denotes the ball centred in
+    Zurich with radius 10km;
+  \item Try to redo what was done in \emph{paper1.pdf};
+  \item Try to redo what was done in \emph{paper2.pdf};
+  \end{itemize}
+\end{slide}
+
 \section{5. Conclusion}
@@
   \begin{itemize}
-  \item Using weighted pairwise likelihood;
   \item Many (many!) utility functions. Highly recommended to have a
@@
   \item Bayesian hierarchical models;
-  \item Unconditional simulations: several implementations (including
-    exact simulations)
   \item Conditional simulations---really CPU demanding. 
@@
   \pause
-  \HandRight{} A rather recent review on max-stable processes with R
-  code is given by \citet{Ribatet2013b}
   \vfill
@@
 
-\begin{wideslide}{References}
-  \vspace*{-1.5em}
-  {\footnotesize
-  \bibliographystyle{apalike}
-  \nobibliography{/Users/Mathieu/Dropbox/biblio_ribatet}
-
-  \begin{itemize}
-  \item[\checkmark] \bibentry{Chandler2007}
-  \item[\checkmark] \bibentry{Cooley2006}
-  \item[\checkmark] \bibentry{Davison2012}
-  \item[\checkmark] \bibentry{Dombry2016}
-  \item[\checkmark] \bibentry{Dombry2017}
-  \item[\checkmark] \bibentry{Padoan2010}
-  \item[\checkmark] \bibentry{Ribatet2013b}
-  \item[\checkmark] \bibentry{Schlather2002}
-  \item[\checkmark] \bibentry{Smith1990}
-  \item[\checkmark] \bibentry{Varin2005}
-  \end{itemize}
-  }
-\end{wideslide}
-
-%\bibliography{/Users/Mathieu/Dropbox/biblio_ribatet}
-%\bibliographystyle{apalike}
-
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/AsAnAside.R

@@ -1,0 +1,13 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Estimate concurrence probabilities w.r.t. Zurich
+idx.station <- 78##78th station corresponds to Zurich
+concurrencemap(rain, coord[,1:2], type = idx.station,
+               plot.border = function(add) swiss(city = TRUE, add = add))
+
+## Estimate the expected area of pointwise concurrence cells (default)
+concurrencemap(rain, coord[,1:2], plot.border = swiss)

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/SpatialTrends.R

@@ -1,0 +1,17 @@
+rm(list=ls())
+library(SpatialExtremes)
+data(rainfall)
+
+## We are going to plot a bunch of circle centered at weather
+## stations, whose radius is proportional to the deviation from the
+## areal mean and whose color is blue or red depending if we're below
+## or above the areal mean
+
+## Symbol plot based on pointwise medians
+symbolplot(rain, coord[,1:2], which = "median", plot.border = swiss,
+           scale = 5, plot.legend = FALSE)
+
+## Symbol plot based on the pointwise GEV estimates (default)
+symbolplot(rain, coord[,1:2], plot.border = swiss,
+           scale = 5, plot.legend = FALSE)
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/pairwiseLlik.R

@@ -1,0 +1,25 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Transform the data to unit Fréchet first
+frechet <- apply(rain, 2, gev2frech, emp = TRUE) ## <<-- we use the empirical CDF
+
+## Fit a bunch of max-stable models using pairwise likelihood
+M0 <- fitmaxstab(frechet, coord[,1:2], "whitmat", nugget = 0)
+M1 <- fitmaxstab(frechet, coord[,1:2], "tpowexp", nugget = 0)
+M2 <- fitmaxstab(frechet, coord[,1:2], "brown")
+
+## Compare the empirical extremal coefficient cloud with fitted extremal coefficient function
+par(mar = c(4, 5, 0.5, 0.5), ps = 14)
+concprob(rain, coord[,1:2], col = "grey")
+concprob(fitted = M0, n.bins = 200, add = TRUE)## <<- you can pass a fitted model also
+plot(M1$conc.prob, from = 0, to = 120, col = 3, add = TRUE)
+plot(M2$conc.prob, from = 0, to = 120, col = 4, add = TRUE)
+legend("topright", c("Schlather", "Extremal-t", "Brown-Resnick"),
+       col = 2:4, lty = 1, bty = "n", inset = 0.05)
+
+
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/simulationFinal.R

@@ -1,0 +1,69 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Define the trend surfaces
+loc.form <- y ~ lon.scaled * lat.scaled
+scale.form <- y ~ lon.scaled + lat.scaled
+shape.form <- y ~ 1
+
+## Center and scale the covariates
+coord.centered <- scale(coord[,1:2], scale = FALSE)
+
+## You can fit the model by invoking the following code
+##
+##M0 <- fitmaxstab(rain, coord[,1:2], "twhitmat", nugget = 0, loc.form,
+##                 scale.form, shape.form, marg.cov = covariates)
+##
+## but for time reasons just do
+load("finalFit.RData")
+
+## Define the grid where we want prediction
+n.grid <- 50
+lon.grid <- seq(645, 770, length = n.grid)
+lat.grid <- seq(209, 291, length = n.grid)
+
+grid <- expand.grid(lon.grid, lat.grid)
+colnames(grid) <- c("lon", "lat") ## <<-- I told you it's important to
+                                  ##      name columns and rows ;-)
+
+## Since we scale the covariates we need to do the same with our grid
+grid.scaled <- grid
+for (i in 1:2)
+    grid.scaled[,i] <- (grid.scaled[,i] - attributes(coord)$"scaled:center"[i])
+
+## Get the prediction for the GEV parameters
+predictions <- predict(M0, grid.scaled)
+loc.pred <- matrix(predictions[,"loc"], n.grid)
+scale.pred <- matrix(predictions[,"scale"], n.grid)
+shape.pred <- matrix(predictions[,"shape"], n.grid)
+
+## Simulate from our fitted model
+
+## First with unit Frechet margins
+
+## For time reasons again...
+##sim <- rmaxstab(1, cbind(lon.grid, lat.grid), "twhitmat",
+##                DoF = M0$param["DoF"], nugget = M0$param["nugget"],
+##                range = M0$param["range"], smooth = M0$param["smooth"],
+##                grid = TRUE)
+load("oneSimFinalModel.RData")
+
+
+## Switch to the appropriate GEV margins
+for (i in 1:n.grid)
+    for (j in 1:n.grid)
+        sim[i,j] <- frech2gev(sim[i,j], loc.pred[i,j], scale.pred[i,j], shape.pred[i,j])
+
+## Plot the realization
[…]

### 2022-03-16 Practicals/practicalSpatialExtremes/Code/spatialGEV.R

@@ -1,0 +1,22 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## First load some data (already formatted)
+data(rainfall)
+
+## Define the trend surfaces
+loc.form <- scale.form <- y ~ lon * lat
+shape.form <- y ~ 1
+
+## Rmk: In the formula above we use lon and lat since these are the
+##      column names of 'coord'
+
+## It is always a good idea to center the "covariates" here
+## coord[,1:2] because the parameters are "more orthogonal"
+coord <- scale(coord, scale = FALSE)
+
+## Compare the two following lines of code and comment
+M0 <- fitspatgev(rain, coord[,1:2], loc.form, scale.form, shape.form)
+M1 <- fitspatgev(rain, scale(coord[,1:2]), loc.form, scale.form, shape.form)
+
+

### 2022-03-16 Practicals/practicalSpatialExtremes/Show/code.R

@@ -1,0 +1,44 @@
+rm(list=ls())
+library(SpatialExtremes)
+
+## Illustration of a max-stable process
+set.seed(1234)
+n.sim <- 50
+coord <- seq(-5, 5, length = 1000)
+Y <- sqrt(2 * pi) * pmax(rgp(n.sim, coord, "powexp", range = 1, smooth = 1.9999), 0)
+poisson <- 1 / cumsum(rexp(n.sim))
+
+Phi <- poisson * Y
+Z <- apply(Phi, 2, max)
+
+par(mar = c(4, 5, 0.5, 0.5), ps = 14)
+matplot(coord, t(Phi), lty = 1, col = "grey", ylab = expression(Z(s)), xlab = expression(s),
+        type = "l")
+lines(coord, Z, col = "orange", lwd = 1.5)
+
+## Illustration of its inner structure
+observed.at <- c(-4, -1, 4)
+idx <- NULL
+for (pos in observed.at)
+    idx <- c(idx, which.min(abs(coord - pos))[1])
+
+idx.extremal.fct <- NULL
+for (j in idx)
+    idx.extremal.fct <- c(idx.extremal.fct, which(Phi[,j] == Z[j]))
+
+plot(coord, Z, col = "orange", lwd = 5,
+     xlab = expression(s), ylab = expression(Z(s)), type = "l",
+     ylim = c(0, max(Phi)))
+
+points(observed.at, Z[idx], pch = 15, col = "seagreen3", cex = 2)
+
+matlines(coord, t(Phi[idx.extremal.fct,]), lty = 1, lwd = 1.5,
+         col = c("red", "black", "blue"))
+
+matlines(coord, t(Phi[-idx.extremal.fct,]), lty = 1, col = "grey")
+
+
+
+
+
+

### 2022-03-16 install.R
Update install.R

@@ -7,6 +7,7 @@ install.packages("extRemes")
 install.packages("fExtremes")
 install.packages("mev")
 install.packages("MCMCpack")
+install.packages("maxLik")
 install.packages("rgl")
 
 

### 2022-03-16 Practicals/multivariate2022.Rmd
test

@@
 par(mar = c(5, 4, 1, 1) + .1)
-WORK = "~/research/enseignement/ATHENS_Cours-Extremes/code/"
-setwd(WORK)
+#WORK = "~/research/enseignement/ATHENS_Cours-Extremes/code/"
+#setwd(WORK)
 options(warn=-1)
@@
 #library for "3D Visualization Using OpenGL":
+options(rgl.useNULL = TRUE)
+options(rgl.printRglwidget = TRUE)
 library(rgl)

### 2022-03-16 Practicals/multivariate2022.Rmd
correction for rgl

@@
 #library for "3D Visualization Using OpenGL":
-options(rgl.useNULL = TRUE)
-options(rgl.printRglwidget = TRUE)
 library(rgl)
@@
 plot3d(w,xlab=expression(w[1]),ylab=expression(w[2]),zlab=expression(w[3]))
+rglwidget()
 
@@
 points3d(0,0,0,col="blue",size=10)
+rglwidget()
 
@@
 points3d(0,0,0,col="blue",size=10)
+rglwidget()
 ```

### 2022-03-19 Datasets/Draft#1.Rmd
draft1

---
title: "Extreme Value Statistics Draft 1"
Dataset: Wind 
Group: 1
Name: Chaitanya Arora, Diandra Gramma 
output: html_document
---

## Introduction ##

Before deep diving into the analysis and coding portion,we should analyse the type of data that we have and what does it mean.

The wind datasets is contains 2 files:

1. "coordinates": This file contains the coordinates of the 5 places in western Europe which we will analyse during the process of the project

2. "dataset": this is a large matrix containing maximum speed in the duration of 3 hours which lasted for more than 3sec. 
The data is from 10th January 1979 - 4th Jan 2014 (35 years)

## Step 1: Basic Analysis of the data ##

1.1 The graphical representation of the 5 places that we are dealing with

```{r}

load("Datasets/1_Wind.RData")
library(ggmap)

bbox <- c(-21.375,  40.875,  16.875,  61.125)

map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")

dataToPlot = data.frame(X = coordinates[, 1], Y = coordinates[, 2])
mapPoints <- ggmap(map) +
  geom_point(
    data = dataToPlot,
    aes(x = X, y = Y),
    color = "red",
    alpha = 1,
    size = 3,
    shape = 16,
  ) +
  labs(title = "", fill = "") +
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
print(mapPoints)

```

### 2022-03-19 Draft#1.Rmd
draft1

---
title: "Extreme Value Statistics Draft 1"
Dataset: Wind 
Group: 1
Name: Chaitanya Arora, Diandra Gramma 
output: html_document
---

## Introduction ##

Before deep diving into the analysis and coding portion,we should analyse the type of data that we have and what does it mean.

The wind dataset is composed of two files:

1. The file named "coordinates": it contains the coordinates of 5 places in north- western Europe, coordinates which we will analyse during the process of the project;

2. The file named "dataset": is a large matrix containing the values of the maximum wind-speed which lasted more than 3 seconds over a time period of 3 hours;

The information found above has been collected over a period of 35 years, more specifically from 10th January 1979 to 4th January 2014.

## Step 1: Basic Analysis of the data ##

1.1 The graphical representation of the 5 places we are analysing in this project

```{r}

load("Datasets/1_Wind.RData")
library(ggmap)

bbox <- c(-21.375,  40.875,  16.875,  61.125)

map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")

dataToPlot = data.frame(X = coordinates[, 1], Y = coordinates[, 2])
mapPoints <- ggmap(map) +
  geom_point(
    data = dataToPlot,
    aes(x = X, y = Y),
    color = "red",
    alpha = 1,
    size = 3,
    shape = 10,
  ) +
  labs(title = "", fill = "") +
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
print(mapPoints)

```
As it can be seen from the graphic above, the 5 places contained in this project can be found in: Dublin, United Kingdom and lastly, the North See. Throughout this report the great saying of Edward Davey "Britain has a lot of wind. It's our wind. We don't have to import it", will be confirmed, as well as a in-depth analysis on how the speed, location and time contribute to it.

1.2 The Basic plots to understand the graph better

The first graphic that will be presented is the histogram. This histogram describes a way of identifying how the speeds of the winds with respect to time influence the frequency of such events happening in all 5 locations.

Therefore, we will have the 5 plots:

```{r}
library(lattice)
data_numeric_v1 <- as.numeric(dataset[,1])
data_numeric_v2 <- as.numeric(dataset[,2])
data_numeric_v3 <- as.numeric(dataset[,3])
data_numeric_v4 <- as.numeric(dataset[,4])
date_numeric_v5 <- as.numeric(dataset[,5])
hist(data_numeric_v1,breaks=30,main="Frequency of wind speeds in the Area v1",xlab = "winds speed in V1")

hist(data_numeric_v2,breaks=30,main="Frequency of wind speeds in the Area v2",xlab = "winds speed in V2")

hist(data_numeric_v3,breaks=30,main="Frequency of wind speeds in the Area v3",xlab = "winds speed in V3")

hist(data_numeric_v4,breaks=30,main="Frequency of wind speeds in the Area v4",xlab = "winds speed in V4")

hist(data_numeric_v4,breaks=30,main="Frequency of wind speeds in the Area v5",xlab = "winds speed in V5")

```

Conclusion:

Despite the fact that the histograms have quite a similar output, they give a rigorous result on the most frequent winds found in the 5 areas.

It can be seen the fact that the most frequent speeds occur between 10-20 km/h, more specifically at about 13/14 km/h. However the figure for the 3rd location seems to represent the fact that for this area, the speed of the wind is frequent at low levels, which only means the fact that this specific point is located in a region with decreased winds.

Another interesting point to be taking into consideration is the fact that the last 2 location have the same exact graphic. (???????? Question, where the hell are these loc and which one is which???????)

The second graphic will be the Empirical distribution one. It will focuse, as the one presented above, on how the speeds of the winds with respect to time influence the frequency of extreme event happening in all of the 5 locations given.

```{r}
dates_new=as.Date(dates)
xyplot(dataset[,1]~dates_new,pch=1,ylab=" ")
xyplot(dataset[,2]~dates_new,pch=1,ylab=" ")
xyplot(dataset[,3]~dates_new,pch=1,ylab=" ")
xyplot(dataset[,4]~dates_new,pch=1,ylab=" ")
xyplot(dataset[,5]~dates_new,pch=1,ylab=" ")

```

<< analysis>>

```{r}
dates_new=as.Date(dates)
yr=factor(year(dates_new),levels=c("1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014"),ordered=TRUE)
bwplot(yr~dataset[,1])
bwplot(yr~dataset[,2])
bwplot(yr~dataset[,3])
bwplot(yr~dataset[,4])
bwplot(yr~dataset[,5])

```
<<< analysis >>

```{r}
dotplot(dataset[,1],pch=1)
dotplot(dataset[,2],pch=1)
dotplot(dataset[,3],pch=1)
dotplot(dataset[,4],pch=1)
dotplot(dataset[,5],pch=1)

```
<<< analysis>>

```{r}
qqmath(dataset[,1])
qqmath(dataset[,2])
qqmath(dataset[,3])
qqmath(dataset[,4])
qqmath(dataset[,5])
```
<< analysis>>

```{r}
qqmath(dataset[,1],distri=qexp)
qqmath(dataset[,2],distri=qexp)
qqmath(dataset[,3],distri=qexp)
qqmath(dataset[,4],distri=qexp)
qqmath(dataset[,5],distri=qexp)
```
<< analysis>>

```{r}
n <- gev.fit(dataset[,1])
m <- gev.diag(n)
``
Doubt: Not able to check the profile liklihood 
since X value is -0.13 it follows this distrubution

### 2022-03-21 Draft#1.Rmd
draft1

@@
 ```{r}
-n <- gev.fit(dataset[,1])
-m <- gev.diag(n)
-``
+var1 <- gev.fit(dataset[,1])
+# Xi = -0.135 --> Weibull
+m <- gev.diag(var1)
+```
+
+```{r}
+var2 <- gev.fit(dataset[,2])
+# Xi = -0.138 --> Weibull 
+n <- gev.diag(var2)
+
+```
+
+```{r}
+var3 <- gev.fit(dataset[,3])
+# Xi = -0.086 --> Gumbell
+o <- gev.diag(var3)
+
+```
+
+```{r}
+var4 <- gev.fit(dataset[,4])
+# Xi = -0.123 --> Weibull
+p <- gev.diag(var4)
+```
+
+```{r}
+var5 <- gev.fit(dataset[,5])
+# Xi = -0.129 --> weibull
+q <- gev.diag(var5)
+```
+
 Doubt: Not able to check the profile liklihood 
 since X value is -0.13 it follows this distrubution
+calculate x n tell distrbution
+
+mean excess plot left
+
+```{r}
+mrl.plot(dataset[,1])
+mrl.plot(dataset[,2])
+mrl.plot(dataset[,3])
+mrl.plot(dataset[,4])
+mrl.plot(dataset[,5])
+```
+
+<<<<<<<<<<<<<<<<<< Univariate Analysis >>>>>>>>>>>>>>>>>>
+
+ANCOVA Analysis (analysis of covariance)
+
+```{r}
+summary(aov(dataset[,1]~dataset[,2]))
+```
+
+Scatterplot
+
+Line of best fit
+
+Regression Analysis
+
+whatever else u find relevant
+
+<<<<<<<<< Multivariate analysis >>>>>
+
+we start with scatterplot
+
+```{r}
+mydata=dataset[,-1]
+pairs(mydata)
+```
+
+ Marginal standardization based on the rank transform
+
+```{r}
+n=nrow(mydata)
+myranks=apply(mydata,2,rank)
+data.unif=myranks/(n+1)
+data.stand=1/(1-data.unif)
+```
+
+Estimation of the extremal coefficient
+
+```{r}
+p=0.8+0:19*0.01
+u.star=1/(1-p)
+u.star[1]*mean(apply(data.stand>u.star[1],1,any))
+u.star[20]*mean(apply(data.stand>u.star[20],1,any))
+theta.est=c()
+for(i in 1:length(p)){
+  theta.est[i]=u.star[i]*mean(apply(data.stand>u.star[i],1,any))
+}
+plot(p,theta.est,type="l",lwd=3,xlab="p",ylab=expression(theta[5]))
+# calculate approximate 95% pointwise confidence intervals
+sd_est = sqrt(u.star/n * theta.est *(1-theta.est/u.star))
+lines(p,theta.est + qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
+lines(p,theta.est - qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
+```
+Estimation of the coefficient of tail dependence
+
+```{r}
+library(fExtremes)
+hillPlot(apply(data.stand,1,min), start=0.8*n, plottype="xi") 
+```

### 2022-03-21 Practicals/multivariate2022.Rmd
draft1

@@
 ```{r, echo=TRUE}
-n=200
+n=500
 pars=c(0.5,0.2,0.2,6)
@@
 
-We first define the log-likelihood function, and then use it within the maxLik function to compute the maximum likelihood estimate.
+We first define the log-likelihood function, and then use it within the maxLink function to compute the maximum likelihood estimate.
 

### 2022-03-21 Draft#1.Rmd
draft2

@@
 ```
+
+Modeling extremal dependence
+
+We first construct the pseudo-polar coordinates and then extract the angles w corresponding to extreme events.
+
+```{r}
+r.obs=apply(data.stand,1,sum)
+w.obs=data.stand/r.obs
+r_0=quantile(r.obs,0.95)
+w.H=w.obs[r.obs>r_0,]
+```
+
+### 3.2) Estimation of joint exceedances above the overall 95% quantile
+
+```{r}
+q95=quantile(mydata,0.95)
+q95
+prob.component=apply(mydata<q95,2,mean)
+u.star=1/(1-prob.component)
+u.star
+ell.min=function(w,u){min(w/u)} #already defined in preceding exercice
+d=5
+n_0=nrow(w.H)
+#estimate the probability of the joint exceedance event:
+d/n_0*sum(apply(w.H,1,ell.min,u=u.star))
+#compare to observed empirical probability:
+mean(apply(t(t(data.stand)/u.star)>1,1,all))
+
+```
+
+Estimation of a parametric Dirichlet model for the angular measure
+
+```{r}
+library(MCMCpack) #for Dirichlet density 
+library(maxLik) #for numerical maximisation of the log-likelihood
+```
+
+```{r}
+
+#define the log-likelihood function:
+llfun=function(alpha){
+if(alpha<=0) return(-Inf) #inadmissible parameters
+ sum(log(ddirichlet(w.H,alpha=rep(alpha,d))))
+}
+
+initpar=1 # define the initial value of the parameter
+fit=maxLik(llfun,start=initpar)
+summary(fit)
+
+```

### 2022-03-24 Draft#1.Rmd
draft_final

--- 2022-03-19 Draft#1.Rmd
@@
 library(ggmap)
+library(lubridate)
 
@@
 ```
-As it can be seen from the graphic above, the 5 places contained in this project can be found in: Dublin, United Kingdom and lastly, the North See.
+As it can be seen from the graphic above, the 5 places contained in this project can be found in: Dublin, United Kingdom and lastly, the North Sea.
 Throughout this report the great saying of Edward Davey "Britain has a lot of wind.
@@
 ```
+Histograms are one of the most used types of plots when trying to plot data, because they are easy to be made.
+In comparison with the real world and the other methods for plotting that will be presented in this report, histograms are not an efficient way of evaluating the data and objectives.
+They are used to create an understanding regarding the discrete or even continuous data measured on a scale.
+
+The clear objective of the report is to find the probability of extreme events happening in years time from present times, however, the extreme events are not to be found clearly and independently in histograms.
+That is why most of the time, these plots are creating an image about the frequency of datasets or even about a historical background on the events happening at a certain time.
 
@@
 Another interesting point to be taking into consideration is the fact that the last 2 location have the same exact graphic.
-(????????
-Question, where the hell are these loc and which one is which???????)
 
@@
 It will focuse, as the one presented above, on how the speeds of the winds with respect to time influence the frequency of extreme event happening in all of the 5 locations given.
+
+As a more in-depth view on the data that we are analysing, in case of the histogram we are measuring the frequency of windspeed in 5 diferent locations (x axis - wind speeds and y axis- frequency).
+
+Based on the figures given, one can grasp the fact that most of the wind speed occur when they are between 10 - 20.
+OF course, some of the locations have the highest frequency on wind speed that are at 22, 25, but all in all they have integrated them in the interval of 10 - 20.
+However, there is one location that has different values to the other ones, more specifically Location 3, where by the positioning of the location, this one is the only one that is on land (some of them are on land, but they are close to the shore).
+Thus, it is affirmative to say that position of the location induces a value for the wind as well as the extreme values that comes along, for which in this point, the highest extreme rates are at 5-6.
+
+<<<<<<<<<< XY Plot >>>>>>>
 
@@
 
-<< analysis>>
+XY are also among the most used plots, along with the histograms.
+By the intuitive name they have, they describe the plot configuration in itself, such that a certain data is allocated and presumed to be compared with the data from the datasets.
+One would mention that these plots are more effectively than histograms, because they do not only show the frequency of events (rare events), but they are also using scattering distribution as a mean access to understanding the data.
+
+Analysis based on the graphs
+
+The data compared is between the 35 years, which were already mentioned, and the intensity of the wind (the wind speed).
+
+The graphics are intuitive and they have a lot to offer in terms of the Extreme values found in certain years.
+
+All of the 5 locations have as a maximum limit the wind speed 30 or 35, whereas the real scattering of the extreme value distribution starts.
+However, two locations’ graphs stand out from all of them: the one for Location 3 and the one for Location 4.
+Now, as it was mentioned at the histogram’s part, it is fully understandable why Location 3 has the lowest values in windspeed, and why from the windspeed of 27/28, the winds throughout the years are considered to be extreme.
+Nevertheless, there are a lot to be mentioned about the intensity with which the points are found, such that in most cases, and this is one way that would work as a pattern, is the fact that the windspeed are in most cases somewhat constant (depending on the region of occurrence).
+
+<< Box Plot >>
 
 ```{r}
-dates_new=as.Date(dates)
+
 yr=factor(year(dates_new),levels=c("1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014"),ordered=TRUE)
@@
 ```
-<<< analysis >>
+In comparison with the above mentioned plot, the box plot is an easier tool and a type of chart often used when data analysis should be conducted.
+They are easy to read, as they summarise data from multiple sources and displays the result just in one graph.
+
+- Displays the data quartiles and the averages.
+- The middle quartile represents the mid-point of the data input, the left-half consists of the values below the middle point and the right-half of the values above the middle point’s value, whilst every “box” is referred to as the inter-quartile range.
+- The whiskers can be lower or upper
+
+Benefits
+
+The box plot is often used, because as in the case of these graphs, one can understand better the statistical impact on a given period of time/ location.
+In regards to the 1-D dot plots, which can easily be made, the information perceived by the user is of higher influence and on a larger scale of detail, which in most cases counts the most.
+
+Analysis based on graphs
+
+To confirm what is written above, in all of the 5 graphs done over the 35 years of analysis, the middle point of the even sized box plots seems to be fluctuating, but not dramatically, meaning that if there would have been taken an average over all the years, this average would not be far from each of the points in all of the years.
+Nonetheless, the box itself depicts which years have been found to have been more drastic on the locations, finally giving a sense of the impact and whether or not the impact was in multiple years, how many years and if it was of high risk or not.
+
+In addition to this, by taking a look at the median points in the box plot, at times there is a discrepancy between a median point from a current year and one from a previous year.
+These points are so called “critical points”, containing a high dataset of values and a low dataset of values.
+Moreover, there can also be depicted at times an almost S-shaped line formed by the middle points of the box plots.
+
+The whiskers are representing the range of our data, thus the whole dataset available is used in this case.
 
@@
 ```
-<<< analysis>>
+The dot plot is usually a type of statistical chart consisting of data contained in the dataset, that is plotted on a single scale.
+It is represented by two axis, the horizontal axis with, the corresponding values and the vertical one that displays the index.
+Thus, in our case the dot plot is a configuration of our wind speed in regards to the time in which it happened, based on all 5 locations.
+
+Analysis based on the graphs
+
+Based on the plots created, one can observe the fact that the general occurrences are between the level of 0 and maximum 30.
+What is here to be considered extremes are the occurrences found after the wind speed of 30.
+
+Moreover, as for the Location 4, which, by the graph it resembles, we can confirm the fact that due to the positioning of this point in Ireland, the winds are to be more powerful.
+The point is right next the the bay leading to North Atlantic Ocean, a part of the Atlantic that is said to be counting high powerful winds.
+However, when taking this point in comparison to Location 1, which is in the North Sea, it can be depicted the fact that occurrences of strong extreme winds are not as often.
+In the fourth location there is a high intercalation of quantiles, whereas for the location in the North Sea, the extreme values are found to be well defined at a point in time.
+
+<<<<QQ Plot>>>
+
+QQ plots, also known as Quantile- Quantile plots and are usually normal when the line formed by the bullet points is as straight as possible.
+They analyse and compare two probability distributions whilst plotting the quantiles between each other.
+If we would have taken the mean and the standard deviation are on the x axis and the random variable on the y axis, we would have received a straight line at an angle of 45 degrees, made to distinguish better the positioning of the actual plot.
+
+QQ plots are used for differentiating ways of plotting and types of distributions, providing at the same time knowledge about locations, scales and type of result the plot provides.
+In addition to this it can provide comparison between datasets, comparison of the distributions, can provide an easily to interpret results and lastly allowing the data to be fit into a distribution.
+
+For the QQ plots a lot of data is necessary, if not the plot won’t lead to a precise result, giving an inconclusive or answer, analyse and type of distribution.
 
@@
 ```
-<< analysis>>
+
+Because the code was written as for a normally distributed QQ plot, one can observe its normality just by looking at the graphics.
+In comparison with the results’ plots done prior, making a QQ normal distribution plot and by interpreting it, one one find in the plot the point that gives a quantile or a precise median.
+One factor needed to be taken into consideration, for example when wanting to  calculate the medians or the quantiles, is always finding the correlation coefficient between paired of sample quantiles.
+The end result, however, would only assume a relative position of the ones mentioned above.
+To generate this, a simple quantiles could be used.
+
+In addition to this, after taking a deeper look on all of the 5 locations’ plots, it can very well be stated that the second location, having the ends of the plot (the scattering ones), can be concluded to be the closes to the Normal distribution.
+However, what is interesting in regards to these plots is the fact that the histogram is of high relation to the way they look:
+
+- If the histogram is a perfectly made parabola, than the normal distribution of the QQ plots on the same data should follow the straight line;
+- If the histogram is also a parabola, but is too peaked in the middle (big discrepancies between the maximum values and the lowest values), the QQ plot will have short distance in common with the straight line and the rest of the first and last quarters would be too below/ too above;
+- If the histogram is skewed to the left, then the normal distribution would also be skewed to the left;
+- If the histogram is skewed to the right, then the normal distribution would also be skewed to the right;
+
+Thus, taking a look back at the histograms done at the first point, it can be confirmed that the normal distribution plots in the ones that have been analysed in this project, that they are asymmetrically to the right, which only means that it is negatively skewed.
 
@@
 ```
-<< analysis>>
+The QQ plots with an exponential distribution are the continuous and independent systems occurring at a more rare constant rate.
+The shape of one is found to be similar with a skewed normal distribution.Because the QQ exponential plot is part of the QQ “family”, for which the theory was presented above, the analysis of the graphics will follow.
+
+One factor needed to be taken into consideration, for example when wanting to calculate the medians or the quantiles, is always finding the correlation coefficient between paired of sample quantiles.
+The end result, however, would only assume a relative position of the ones mentioned above.
+To generate this, a simple quantiles could be used.
+
+The process, as in the other cases, was realized based on data we have available, the windspeed, years and finally the location.
+The graphs seem to be similar from one to another in a curved shape.
+However, the most important factor is the scattering dots and how they occur.
+Based on this factor, Location 3 is the one that is the most linear dependent out of the rest, following the imaginary (or not) straight line that has the origin in 0.
+The rest of the locations seem to be independent, rare and extreme events.
+
+Despite this, in exponential distribution, one finds the graph closest to the straight line the best fit, whereas in our case, as already mentioned is not the case.
+This is due to the parameter lambda, p, t and x, where lambda must always have a positive value, p that find values between 0 and 1, x is also always positive and the threshold t is higher than x.
+
+In the case analysed, there is a discrepancy in all locations between the values: for windspeed (0-20) and for the exponential quantiles (0-2 in some cases and 0-4 in other cases), leading to the almost wrongful graphics conducted, after running the code.
+
+Finally, if an more in-depth process can be done, more so seeing just the quantiles for the exponential of the tail, then, as for the last case, the Location 3 is considered to be the best fit, whereas Location 5 is the least suitable for this type of plots, however non of the graphics are linearly related.
+
+<<<< GEV Plots >>>
+
+GEV plots also known as Generalised Extreme Value Distribution is a type of modelling that make use of the maximum likelihood estimation.
+
+For simulating the block maximum data, The GEV can be characterised valuably as the constraining dispersion of square maxima (or minima). That's , in case you create an expansive number of free arbitrary values from a single likelihood dispersion, and take their greatest esteem, the dissemination of that most extreme is roughly a GEV.
+The unique conveyance decides the shape parameter, k, of the coming about GEV dissemination. Conveyances whose tails drop off as a polynomial, such as Student's t, lead to a positive shape parameter. Dispersions whose tails diminish exponentially, such as the typical, compare to a zero shape parameter. Conveyances with limited tails, such as the beta, compare to a negative shape parameter.
+
+Two of the graphs shown in the plots have already been analysed and commented on prior, thus the GEV plot analysis will focus on the other two left: the Probability Plot and the Return Level Plot.
+
+The first one mentioned, as the name imposes “probability”, resembles the process of assessing your data from the present datasets with different distributions, acknowledging whether or not it resembles one (an example for this would be the normal distribution).
+It is formed by two axis, more specifically x which is the order statistic median for the given distribution and on the y axis, the ordered response values.
+Moreover, the plot should follow a distribution of points that are following a straight line.
+
+Regarding the statement provided above and taking into consideration the graphics made for al 5 locations, one can suggest that the most accurate graphics between all of them is both the first location and the fourth one.
+
+Talking about return level plots, there is to be noticed a fine simplicity of interpretation and a good operation when the graphic is also validate (this is usually due to the tail compression of the distribution when in extrapolation).
+
+An analysis done on the threshold of the graphic can be found in the following way: if the threshold is too low, the asimptotic model is invalid, if the threshold is too high, then there is variance is too large due to some of the data points.
 
 ```{r}
-n <- gev.fit(dataset[,1])
-m <- gev.diag(n)
-``
-Doubt: Not able to check the profile liklihood 
-since X value is -0.13 it follows this distrubution
+library(ismev)
+var1 <- gev.fit(dataset[,1])
+# Xi = -0.135 --> Weibull
+m <- gev.diag(var1)
+```
+
+```{r}
+var2 <- gev.fit(dataset[,2])
+# Xi = -0.138 --> Weibull 
+n <- gev.diag(var2)
+
+```
+
+```{r}
+var3 <- gev.fit(dataset[,3])
+# Xi = -0.086 --> Gumbell
+o <- gev.diag(var3)
+
+```
+
+```{r}
+var4 <- gev.fit(dataset[,4])
+# Xi = -0.123 --> Weibull
+p <- gev.diag(var4)
+```
+```{r}
+var5 <- gev.fit(dataset[,5])
+# Xi = -0.129 --> weibull
+q <- gev.diag(var5)
+```
+
+<<<<<Mean excess plot>>>>
+
+As with the other plots, the Mean Excess Plot is one that is also used in extreme values statistics, and used to aid the variability of pick when it comes to the threshold of the process.
+Moreover is a way of collecting information about the adequacy of the GPD model inpractice.
+
+But what is the GPD model?
+It is a type of distribution that has a very similar work domain as the one of central limit theorem.
+It focuses on the maximum values of a random variable, which is independent and identically distributed.
+As in for our case with the Mean Excess plot, the GPD model focuses the most on the right tail part of the the distribution, thus only the positive extreme values.
+
+The Mean Excess plot is also a tool to be compared with other techniques, such as the Hill estimator, the Pickands estimator and the QQ plot.
+
+Analysis based on the graph
+
+In other words, the way it works is by conditioning a random variable X on the event X>t, where the mean-excess function, also known as mean residual life in survival analysis is e(t)=E[X-t | X>t], where X is a random variable.
+This analysis focuses on the ratio between the mean excess function and the windspeed data from the datasets.
+Moreover, the plots were done for all of the 5 locations.
+
+By just looking at the graphic, one can state the fact that the type of plot is underlined by a thin tailed behaviour, because it is a plot from the top left continuing to the right bottom of the graphic (downwards slope).
+In addition to this, it can also be mentioned the fact that the graphics are constant leaning towards an exponential distribution.
+
+Regarding the right tail behaviour of the plots, one can assume that if the tail tends to a straight line with positive, negative or zero slopes, then the analysis is in the domain of attraction a Fréchet, Gumbel or Weibull distributions.
+And such is in this case, a Generalised Pareto with random variables α = 2 and β = 2.
+
+```{r}
+mrl.plot(dataset[,1])
+mrl.plot(dataset[,2])
+mrl.plot(dataset[,3])
+mrl.plot(dataset[,4])
+mrl.plot(dataset[,5])
+```
+
+<<<<<<<<<<<<<<<<< bi-variate Analysis >>>>>>>>>>>>>>>>>>
+
+bi-variarte analysis is used to understand the relation of data's one attrubute to the other attrubute, for this we will try to use the line of best fit model which is used a lot in regression analysis and later scatter plot to understand the bivariate relation ship
+
+This is the line of best fit
+```{r}
+dates_new=as.Date(dates)
+plot(dataset[,2],dataset[,1])
+abline(lm(dataset[,2]~dataset[,1]))
+
+```
+Since it does not yield anything fruitful due to a wide range of values throughout the graph, therfore we will avoid doing it for other ones as well.
+
+<<<<<<<<< Multivariate analysis >>>>>>>>
+
+Before beginning multivariate analysis lets talk about what is multivariate anlaysis what is the approach we will be taking up and why?
+
+A multivariate distribution is when we have multiple iid (independent identical distrubution) vectors.
+One point to note is that there is no natural ”ordering” or unique definition of extremes in a multivariate setting.
+
+Xi = (Xi,1,...,Xi,D) ∈ R^d.
+
+Extremes in terms of Multivariate distribution is given in different forms by barnett is 
+1. Component wise maxima 
+2. Convex hull
+3. Structure variable
+
+The third one is used very often in different fields namely portfolio risk management or calculating wind speed etc.
+
+In multivariate analysis of extreme value using structural approach we calculate the joint distribution F which is given by
+
+Pr(X ∈ A) = Integral(dF(x))
+
+for some set A ⊂ RD .
+
+The problem is the fact that we have little or no data to predict if there will be an event which will occcur a certain value 
+Therefore we use asymptotically motivated models to extrapolate into the multivariate tail.
+
+In order to do that we encounter various problems like that each point can be extraploted in many directions and Multiple models are available with different hypothesis.
+
+Therefore we can use the Copulas to model dependence structure which is unique and determines the dependent structure of X. The marginal choice varies from distribution to distribution
+
+1. Extremal Independence:
+
+Extremal independence means that what is the probablity of a value occurring over a certian value.
+
+pairwise extermal independence is calculated with (Xi) coefficient r(Y1,Y2)
+
+# Pr{Y1 >u|Y2 >u}= uPr{Y1 >u,Y2 >u}/uPr{Y2 > u}
+
+The value of Xi is used to calculate asymptotic external dependence 
+If the Y1 and Y2 are independent then it is 0
+
+Asymptotic independence is calculated with the use of η (ita)
+
+Now that we have established a good understanding of the multivariate analysis and the different approaches used:
+
+<<<< Scatter plot diagram :>>>
+
+```{r}
+mydata=dataset[,-1]
+pairs(mydata)
+```
+1.Scatter plot is a graph in which the values of two variables are plotted along two axes, the pattern of the resulting points revealing any correlation present.
+
+2.We are using this plot just to understand the basic correlation between the wind speed in different areas to see if there is a relation between them (Does the wind speed in one place affect the other)
+
+3.Looking at the graph below we observe: Since each and every graph shows a positive correaation that signifies that 
+There is a positive correaltiion between the two variables
+
+4. This will be relevant for us to to understand that there is some proportional realtionship between the windspeed in different areas.
+
+<<<<Marginal standardization based on the rank transform>>>
+
+```{r}
+n=nrow(mydata)
+myranks=apply(mydata,2,rank)
+data.unif=myranks/(n+1)
+data.stand=1/(1-data.unif)
+```
+
+The second graph in the multivariate analysis is Marginal Standardization Distribution.
+Marginal Standardization graph is in which the values of two variables are plotted along two axes, the pattern of the resulting points revealing any correlation present.
+
+This can be done using various models depending upon the type of distribution:
+Continuous we use: Fr ́echet marginal distribution 
+or we can also use the standard parto distribution
+
+<<<<Estimation of the extremal coefficient>>>>>
+
+```{r}
+p=0.8+0:19*0.01
+u.star=1/(1-p)
+u.star[1]*mean(apply(data.stand>u.star[1],1,any))
+u.star[20]*mean(apply(data.stand>u.star[20],1,any))
+theta.est=c()
+for(i in 1:length(p)){
+  theta.est[i]=u.star[i]*mean(apply(data.stand>u.star[i],1,any))
+}
+plot(p,theta.est,type="l",lwd=3,xlab="p",ylab=expression(theta[5]))
+# calculate approximate 95% pointwise confidence intervals
+sd_est = sqrt(u.star/n * theta.est *(1-theta.est/u.star))
+lines(p,theta.est + qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
+lines(p,theta.est - qnorm(.975)*sd_est, col = "blue", lwd = 3, lty = 2)
+```
+1. Extermal coefficeint is a natural way to measure dependence among spatial maxima stems from considering the distribution of the largest value.
+
+2. We would be using this graph to get the probability of events which are happening above a certain extent i.e. the probability of an extreme event occurring which will be above a threshold
+
+3. This graph is intetprated as theta belongs to [1,d] is the avg number of independent clusters among the components 1 to d where an extreme event occurs
+
+4. Judging from the significantly increasing estimates when p increases, the data do not show threshold-stability. We seem to be in the situation of asymptotic independence.
+
+<<<<Estimation of the coefficient of tail dependence>>>>
+
+```{r}
+library(fExtremes)
+hillPlot(apply(data.stand,1,min), start=0.8*n, plottype="xi") 
+```
+The concept of tail dependence describes the amount of dependence in the lower- left-quadrant tail or upper-right-quadrant tail of a multivariate distribution.
+A common measure of tail dependence is given by the so-called tail-dependence coefficient.
+
+This is estimated using the Hill estimator applied to the sample of the minimum of the 5 components which is the five locations in this case.
+This can be done conveniently with the hillPlot function in the fExtremes package, which also provides confidence intervals.
+
+As seen from the graph the hillplot graph shoes a negative slope for the in increasing values, and the value of ita is less than one and is constantly decreasing which indicates a strong asympototic independence between the 5 vairables together.
+
+Which means that at the extreme values the windspeed at different places do not seem to affect the other places.
+
+<<<<<Modeling extremal dependence:>>>>>>>
+
+```{r}
+r.obs=apply(data.stand,1,sum)
+w.obs=data.stand/r.obs
+r_0=quantile(r.obs,0.95)
+w.H=w.obs[r.obs>r_0,]
+```
+
+```{r}
+q95=quantile(mydata,0.95)
+q95
+prob.component=apply(mydata<q95,2,mean)
+u.star=1/(1-prob.component)
+u.star
+ell.min=function(w,u){min(w/u)} #already defined in preceding exercice
+d=5
+n_0=nrow(w.H)
+#estimate the probability of the joint exceedance event:
+d/n_0*sum(apply(w.H,1,ell.min,u=u.star))
+#compare to observed empirical probability:
+mean(apply(t(t(data.stand)/u.star)>1,1,all))
+
+```
+
+Moving on right now even though we are referring to the extreme events and analyzing them with respect to the other elements, but majorly we are referring to the fact if there is a connection between them to make preditions and calculated risks according to the data
+
+But now we can use these asymptomatic dependent models to provide and idea for conservative estimates for the joint risks for this we make use of Modeling extremal dependence
+
+To accomplish this we construct the pseudo-polar coordinates and then extract the angles w corresponding to extreme events, after which we filter that data to the extreme values by taking joint exceedances above the overall 95% quantile
+
+Looking at the data we can see that empirical estimate we see that the estimate is relatively close to the empirically observed value.
+
+<<<<<<<<Conclusions>>>
+
+Q1. How confidence are scientists when having such a limited amount of knowledge when trying to esteem the effects on the long term?
+
+In these cases, the data that we are using are based on extrapolation.
+The data is meant to be working efficiently in providing an answer for the future, as in all of the Extreme Values Statistics.
+The objective is highly defined as providing a probability statement on the events that are to be/ or not to be happening in the next far future.
+However, In most cases, the data that we detain, regardless of the fact that is providing numerous elements of information or not, could not include probability affirmations for the future.
+Everything, in the end is just an ambigue estimation.
+
+Thus, the scientist do not have something concrete to assume in the future, they are basing themselves on extrapolating concrete data from the past and present with which therefore are constructing the base for the future.
+By extrapolating the data that they had, in the past, they assumed moments in which the aftermath was assumed to be corect.
+
+Finally, even though the “uncertainty” is present, the best fit for this objective is comparing models, divide the distribution in quantile, examine at a larger scale and at a more in-depth scale and compare the quantiles between the models.
+This is a way of finding the best fit for models, whether they are relating to finance (stock market, inflation …) or natural phenomena (wind, rain, heat waves …).
+
+Q2. Would changing the data, when evaluating methods, would impact the process itself? Would it still perform well?
+
+To some extent yes, As we know that when we get data there are always some boiler plate stuff that we should conduct which will help us in understanding the data more concretely, for example when we had the wind data we:
+
+Firstly looked over the data and tried to understand what kind of attributes each file has and what does it mean.
+
+Secondly we took the data and we should have cleaned it as the raw data that we get is not clean is not a structured form: this process is called data cleaning. 
+Ps. we skipped this step as we got it normally from the professor and was not exactly raw data
+
+Third task is to deepen the understanding of the data, here is where we make use of basic plots to go deeper, we did this by making use of histograms,xy plots, dot plots etc.
+They helped us in understanding where does gnerally data lie and more importantly what does it mean
+
+In our opinion the 4th task is to undertand the patterns in the data and picking up a topic or a idea which we want to know about, this can be formed before you start the project or while doing as well as per the interest.
+then the analysis starts getting a little specific depending upon the ques we are trying to answer.
+
+For eg.
+after doing the plots and understanding what distribution they are following I want to explore the relation or impact of extreme wind speeds in one place and if affects the other or How much are they related together etc.
+etc.
+TO answer each of these specific ques one specific ques needs to be done
+
+To sum it up most of the process to develop a deep understanding of the data remains same but as we go further the process become specific to the ques you are trying to answer.

### 2022-03-24 Draft#1.Rmd
draft_changes

@@
 ---
-title: "Extreme Value Statistics Draft 1"
+Title: "Extreme Value Statistics Draft 1"
 Dataset: Wind 
@@
 
-## Introduction ##
+################# Introduction ####################
 
@@
 
-## Step 1: Basic Analysis of the data ##
+<<<<<<<<<. Basic Analysis of the data >>>>>>>>>>>>
 
-1.1 The graphical representation of the 5 places we are analysing in this project
+1 The graphical representation of the 5 places we are analysing in this project
 
@@
 
-1.2 The Basic plots to understand the graph better
+2. The Basic plots to understand the graph better
 
@@
 
-<<<<<<<<<< XY Plot >>>>>>>
+3. <<<<<<<<<< XY Plot >>>>>>>
 
@@
 
-<< Box Plot >>
+4. <<<<< Box Plot >>>>>
 
@@
 
+5. <<<<<< Dot plots >>>>>>
+
 ```{r}
@@
 
-<<<<QQ Plot>>>
+6. <<<<<< QQ Plot >>>>>
 
@@
 
-<<<< GEV Plots >>>
+7. <<<<<< GEV Plots >>>>>>
 
@@
 
-<<<<<Mean excess plot>>>>
+8. <<<<<<<<Mean excess plot>>>>>>
 
@@
 
-<<<<<<<<<<<<<<<<< bi-variate Analysis >>>>>>>>>>>>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< bi-variate Analysis >>>>>>>>>>>>>>>>>>
 
@@
 This is the line of best fit
+
+9. <<<<<Line of best fit>>>>>
+
 ```{r}
@@
 
-<<<<<<<<< Multivariate analysis >>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< Multivariate analysis >>>>>>>>
 
@@
 
-1. Extremal Independence:
+Extremal Independence:
 
@@
 
-<<<< Scatter plot diagram :>>>
+10. <<<<<<<<<< Scatter plot diagram :>>>>>>>>>>
 
@@
 
-<<<<Marginal standardization based on the rank transform>>>
+11. <<<<<<<<<< Marginal standardization based on the rank transform >>>>>>>>
 
@@
 
-<<<<Estimation of the extremal coefficient>>>>>
+12. <<<<<<<<< Estimation of the extremal coefficient >>>>>>>>>>>
 
@@
 
-<<<<Estimation of the coefficient of tail dependence>>>>
+13. <<<<<<<<< Estimation of the coefficient of tail dependence >>>>>>
 
@@
 
-<<<<<Modeling extremal dependence:>>>>>>>
+14. <<<<< Modeling extremal dependence: >>>>>>>
 
@@
 
-<<<<<<<<Conclusions>>>
+15. <<<<<<<< Conclusions >>>>>>>
 
